### 2023-06-25 DHI_GHI_2_sza_trends.tex
Commit 2023-06-25 21:17

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Trends of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-06-25}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI and DNI radiation for `clear sky' and all sly conditions.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
fANCOVA 0.6-1 loaded
\end{verbatim}

\begin{verbatim}
Loading required package: zoo
\end{verbatim}

\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}

Load environment and data from:  ~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda 
\end{verbatim}

\hypertarget{long-term-by-sza}{%
\subsection{2. Long term by SZA}\label{long-term-by-sza}}

\hypertarget{data-info}{%
\subsubsection{Data info}\label{data-info}}

Time data span 1993-04-12, 2023-05-31

\hypertarget{process}{%
\subsubsection{Process}\label{process}}

\FloatBarrier

\hypertarget{plot-of-sza-trends}{%
\subsection{Plot of SZA trends}\label{plot-of-sza-trends}}

\newpage

\hypertarget{all-sky-cond.-dir.-beam-irrad.}{%
\paragraph{All sky cond. Dir. Beam Irrad.}\label{all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-4} \end{center}

\newpage

\hypertarget{clear-sky-cond.-dir.-beam-irrad.}{%
\paragraph{Clear sky cond. Dir. Beam Irrad.}\label{clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-8} \end{center}

\newpage

\hypertarget{cloudy-cond.-dir.-beam-irrad.}{%
\paragraph{Cloudy cond. Dir. Beam Irrad.}\label{cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-12} \end{center}

\newpage

\hypertarget{all-sky-cond.-sdr}{%
\paragraph{All sky cond. SDR}\label{all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-16} \end{center}

\newpage

\hypertarget{clear-sky-cond.-sdr}{%
\paragraph{Clear sky cond. SDR}\label{clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-20} \end{center}

\newpage

\hypertarget{cloudy-cond.-sdr}{%
\paragraph{Cloudy cond. SDR}\label{cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-24} \end{center}

\hypertarget{plot-of-sza-trends-for-each-season-of-year}{%
\subsection{Plot of SZA trends for each season of year}\label{plot-of-sza-trends-for-each-season-of-year}}

\newpage

\hypertarget{winter-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter All sky cond. Dir. Beam Irrad.}\label{winter-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-4} \end{center}

\newpage

\hypertarget{winter-all-sky-cond.-sdr}{%
\subsubsection{Winter All sky cond. SDR}\label{winter-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-8} \end{center}

\newpage

\hypertarget{winter-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter Clear sky cond. Dir. Beam Irrad.}\label{winter-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-12} \end{center}

\newpage

\hypertarget{winter-clear-sky-cond.-sdr}{%
\subsubsection{Winter Clear sky cond. SDR}\label{winter-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-16} \end{center}

\newpage

\hypertarget{winter-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter Cloudy cond. Dir. Beam Irrad.}\label{winter-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-20} \end{center}

\newpage

\hypertarget{winter-cloudy-cond.-sdr}{%
\subsubsection{Winter Cloudy cond. SDR}\label{winter-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-24} \end{center}

\newpage

\hypertarget{spring-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring All sky cond. Dir. Beam Irrad.}\label{spring-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-28} \end{center}

\newpage

\hypertarget{spring-all-sky-cond.-sdr}{%
\subsubsection{Spring All sky cond. SDR}\label{spring-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-32} \end{center}

\newpage

\hypertarget{spring-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring Clear sky cond. Dir. Beam Irrad.}\label{spring-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-36} \end{center}

\newpage

\hypertarget{spring-clear-sky-cond.-sdr}{%
\subsubsection{Spring Clear sky cond. SDR}\label{spring-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-38} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-40} \end{center}

\newpage

\hypertarget{spring-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring Cloudy cond. Dir. Beam Irrad.}\label{spring-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-44} \end{center}

\newpage

\hypertarget{spring-cloudy-cond.-sdr}{%
\subsubsection{Spring Cloudy cond. SDR}\label{spring-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-46} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-47} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-48} \end{center}

\newpage

\hypertarget{summer-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer All sky cond. Dir. Beam Irrad.}\label{summer-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-49} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-50} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-51} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-52} \end{center}

\newpage

\hypertarget{summer-all-sky-cond.-sdr}{%
\subsubsection{Summer All sky cond. SDR}\label{summer-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-53} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-54} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-55} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-56} \end{center}

\newpage

\hypertarget{summer-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer Clear sky cond. Dir. Beam Irrad.}\label{summer-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-57} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-58} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-59} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-60} \end{center}

\newpage

\hypertarget{summer-clear-sky-cond.-sdr}{%
\subsubsection{Summer Clear sky cond. SDR}\label{summer-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-61} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-62} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-63} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-64} \end{center}

\newpage

\hypertarget{summer-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer Cloudy cond. Dir. Beam Irrad.}\label{summer-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-65} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-66} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-67} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-68} \end{center}

\newpage

\hypertarget{summer-cloudy-cond.-sdr}{%
\subsubsection{Summer Cloudy cond. SDR}\label{summer-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-69} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-70} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-71} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-72} \end{center}

\newpage

\hypertarget{autumn-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn All sky cond. Dir. Beam Irrad.}\label{autumn-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-73} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-74} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-75} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-76} \end{center}

\newpage

\hypertarget{autumn-all-sky-cond.-sdr}{%
\subsubsection{Autumn All sky cond. SDR}\label{autumn-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-77} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-78} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-79} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-80} \end{center}

\newpage

\hypertarget{autumn-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn Clear sky cond. Dir. Beam Irrad.}\label{autumn-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-81} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-82} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-83} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-84} \end{center}

\newpage

\hypertarget{autumn-clear-sky-cond.-sdr}{%
\subsubsection{Autumn Clear sky cond. SDR}\label{autumn-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-85} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-86} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-87} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-88} \end{center}

\newpage

\hypertarget{autumn-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn Cloudy cond. Dir. Beam Irrad.}\label{autumn-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-89} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-90} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-91} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-92} \end{center}

\newpage

\hypertarget{autumn-cloudy-cond.-sdr}{%
\subsubsection{Autumn Cloudy cond. SDR}\label{autumn-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-93} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-94} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-95} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-96} \end{center}

\textbf{END}

\begin{verbatim}
2023-06-25 17:38:40.8 athan@tyler DHI_GHI_1_longterm_trends.R 1.213719 mins
\end{verbatim}

\end{document}