### 2023-06-25 DHI_GHI_3_trends_consistency.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Trends of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-06-26}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI and DNI radiation for `clear sky' and all sly conditions.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
fANCOVA 0.6-1 loaded
\end{verbatim}

\begin{verbatim}
Loading required package: zoo
\end{verbatim}

\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}

Load environment and data from:  ~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda 
\end{verbatim}

\hypertarget{consistency-of-trends}{%
\subsection{3. Consistency of trends}\label{consistency-of-trends}}

\hypertarget{data-info}{%
\subsubsection{Data info}\label{data-info}}

Time data span 1993-04-12, 2011-03-31

Where is a \textbf{running mean the window is 1826.2125 days} or
5 years.

\hypertarget{process}{%
\subsubsection{Process}\label{process}}

\newpage
\FloatBarrier

\hypertarget{whole-day-daily-cumulative-sum}{%
\subsubsection{Whole day daily cumulative sum}\label{whole-day-daily-cumulative-sum}}

All data in this section are referring to daily means.

\newpage

\hypertarget{daily-cum-sums-for-all-sky-cond.}{%
\paragraph{Daily cum sums for All sky cond.}\label{daily-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 2610 & -3026 & 18 & -31.52 & -415.8 \\
\textbf{1994} & 5058 & -5349 & 24.79 & -35.9 & -291.5 \\
\textbf{1995} & 3670 & -5098 & 22.38 & -29.47 & -1429 \\
\textbf{1996} & 3269 & -7514 & 19.23 & -41.06 & -4245 \\
\textbf{1997} & 5002 & -5946 & 24.88 & -38.87 & -944.7 \\
\textbf{1998} & 4977 & -5880 & 23.59 & -42.3 & -902.3 \\
\textbf{1999} & 5124 & -4687 & 23.94 & -31.67 & 437.1 \\
\textbf{2000} & 6492 & -3419 & 25.97 & -32.25 & 3073 \\
\textbf{2001} & 4913 & -5082 & 23.97 & -36.56 & -168.3 \\
\textbf{2002} & 5152 & -5694 & 26.56 & -34.93 & -542.2 \\
\textbf{2003} & 5420 & -5678 & 25.93 & -37.85 & -257.8 \\
\textbf{2004} & 5095 & -5125 & 23.81 & -36.09 & -29.19 \\
\textbf{2005} & 5160 & -3690 & 24.93 & -29.76 & 1469 \\
\textbf{2006} & 5554 & -4210 & 26.2 & -34.22 & 1344 \\
\textbf{2007} & 4938 & -3920 & 23.74 & -35.96 & 1019 \\
\textbf{2008} & 5814 & -4603 & 24.23 & -39.01 & 1212 \\
\textbf{2009} & 5724 & -4746 & 28.06 & -34.39 & 977.8 \\
\textbf{2010} & 4347 & -4191 & 26.84 & -34.92 & 156.5 \\
\textbf{2011} & 1707 & -2171 & 40.65 & -48.24 & -463.5 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-4} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-clear-sky-cond.}{%
\paragraph{Daily cum sums for Clear sky cond.}\label{daily-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 3054 & -3389 & 22.13 & -33.56 & -334.9 \\
\textbf{1994} & 6247 & -5727 & 31.24 & -37.93 & 520.4 \\
\textbf{1995} & 4210 & -5814 & 28.45 & -30.93 & -1604 \\
\textbf{1996} & 4206 & -8033 & 26.62 & -41.19 & -3827 \\
\textbf{1997} & 5268 & -6573 & 29.43 & -37.56 & -1304 \\
\textbf{1998} & 5243 & -6363 & 28.34 & -38.56 & -1119 \\
\textbf{1999} & 5920 & -5190 & 28.88 & -33.05 & 730.8 \\
\textbf{2000} & 6699 & -3870 & 29.64 & -29.77 & 2829 \\
\textbf{2001} & 4720 & -5592 & 26.08 & -35.17 & -872.5 \\
\textbf{2002} & 5795 & -6316 & 32.56 & -35.68 & -520.3 \\
\textbf{2003} & 5721 & -6286 & 30.11 & -37.19 & -565.2 \\
\textbf{2004} & 5192 & -5678 & 28.84 & -32.26 & -485.9 \\
\textbf{2005} & 6115 & -4224 & 32.7 & -29.54 & 1891 \\
\textbf{2006} & 6389 & -4608 & 32.43 & -33.63 & 1781 \\
\textbf{2007} & 5666 & -4235 & 29.21 & -34.43 & 1432 \\
\textbf{2008} & 6599 & -4942 & 29.59 & -36.88 & 1656 \\
\textbf{2009} & 6003 & -5238 & 32.28 & -33.58 & 765.2 \\
\textbf{2010} & 4117 & -4603 & 28.01 & -34.1 & -486.4 \\
\textbf{2011} & 1740 & -2226 & 42.43 & -49.46 & -486.1 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-8} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-cloudy-cond.}{%
\paragraph{Daily cum sums for Cloudy cond.}\label{daily-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 556.6 & -1021 & 21.41 & -23.76 & -464.8 \\
\textbf{1994} & 616.9 & -1544 & 17.14 & -22.7 & -926.6 \\
\textbf{1995} & 698 & -899.1 & 17.9 & -21.41 & -201.1 \\
\textbf{1996} & 957.9 & -1481 & 30.9 & -24.28 & -523.2 \\
\textbf{1997} & 1305 & -996.4 & 25.59 & -24.3 & 308.4 \\
\textbf{1998} & 1227 & -761.8 & 24.06 & -18.14 & 465 \\
\textbf{1999} & 754.4 & -1116 & 16.4 & -18.6 & -361.7 \\
\textbf{2000} & 1015 & -993.2 & 22.55 & -21.59 & 21.44 \\
\textbf{2001} & 1047 & -576 & 26.17 & -16.46 & 470.8 \\
\textbf{2002} & 1282 & -1019 & 23.73 & -21.67 & 263 \\
\textbf{2003} & 1123 & -1041 & 26.13 & -21.69 & 82.5 \\
\textbf{2004} & 1395 & -562.8 & 27.9 & -19.41 & 832.3 \\
\textbf{2005} & 904.3 & -1018 & 18.84 & -22.13 & -113.7 \\
\textbf{2006} & 1044 & -929.3 & 21.75 & -23.23 & 114.7 \\
\textbf{2007} & 701.4 & -1082 & 25.05 & -23.03 & -380.8 \\
\textbf{2008} & 916.9 & -1225 & 22.36 & -21.13 & -308.6 \\
\textbf{2009} & 894.8 & -575.7 & 22.37 & -19.85 & 319.1 \\
\textbf{2010} & 790.5 & -469.3 & 23.25 & -19.55 & 321.2 \\
\textbf{2011} & 96.34 & -14.4 & 16.06 & -3.6 & 81.94 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{whole-day-monthly-cumulative-sum}{%
\subsubsection{Whole day monthly cumulative sum}\label{whole-day-monthly-cumulative-sum}}

In this section the monthly values have been calculated from daily mean values.

\newpage

\hypertarget{monthly-cum-sums-for-all-sky-cond.}{%
\paragraph{Monthly cum sums for All sky cond.}\label{monthly-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1392}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 32.03 & -43.15 & 8.008 & -10.79 & -11.12 \\
\textbf{1994} & 38.66 & -59.06 & 6.443 & -9.843 & -20.4 \\
\textbf{1995} & 33.47 & -77.73 & 6.694 & -12.96 & -44.26 \\
\textbf{1996} & 6.869 & -161.9 & 2.29 & -17.99 & -155.1 \\
\textbf{1997} & 33.28 & -61.31 & 8.321 & -7.664 & -28.02 \\
\textbf{1998} & 37.8 & -58.98 & 4.726 & -14.74 & -21.17 \\
\textbf{1999} & 27.94 & -15.43 & 3.992 & -3.086 & 12.51 \\
\textbf{2000} & 101.9 & -10.79 & 9.266 & -10.79 & 91.13 \\
\textbf{2001} & 60.36 & -61.26 & 10.06 & -10.21 & -0.9042 \\
\textbf{2002} & 51.82 & -69.79 & 12.95 & -8.724 & -17.98 \\
\textbf{2003} & 35.53 & -34.82 & 7.106 & -4.974 & 0.7085 \\
\textbf{2004} & 43.12 & -43.88 & 5.39 & -10.97 & -0.7636 \\
\textbf{2005} & 55.88 & -8.827 & 7.983 & -2.207 & 47.06 \\
\textbf{2006} & 50.15 & -25.46 & 7.165 & -6.365 & 24.69 \\
\textbf{2007} & 39.78 & -8.71 & 6.63 & -2.177 & 31.07 \\
\textbf{2008} & 73.43 & -25.38 & 9.179 & -6.345 & 48.05 \\
\textbf{2009} & 66.67 & -38.18 & 9.524 & -7.635 & 28.49 \\
\textbf{2010} & 32.42 & -28.63 & 6.484 & -9.544 & 3.789 \\
\textbf{2011} & 5.498 & -19.1 & 5.498 & -9.55 & -13.6 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-4} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-clear-sky-cond.}{%
\paragraph{Monthly cum sums for Clear sky cond.}\label{monthly-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 33.26 & -39 & 11.09 & -7.8 & -5.739 \\
\textbf{1994} & 54.63 & -46.3 & 9.105 & -7.716 & 8.331 \\
\textbf{1995} & 42.76 & -87.74 & 7.127 & -14.62 & -44.98 \\
\textbf{1996} & 17.19 & -158.6 & 3.437 & -22.66 & -141.5 \\
\textbf{1997} & 32.75 & -74.28 & 8.188 & -9.285 & -41.52 \\
\textbf{1998} & 32.38 & -62.16 & 4.625 & -12.43 & -29.79 \\
\textbf{1999} & 35.78 & -14.45 & 4.473 & -3.612 & 21.34 \\
\textbf{2000} & 92.01 & -10.09 & 8.365 & -10.09 & 81.92 \\
\textbf{2001} & 53.16 & -82.67 & 8.86 & -13.78 & -29.51 \\
\textbf{2002} & 54.62 & -72.14 & 13.65 & -9.017 & -17.52 \\
\textbf{2003} & 28.23 & -37.42 & 7.057 & -4.678 & -9.193 \\
\textbf{2004} & 33.27 & -48.89 & 5.545 & -8.149 & -15.62 \\
\textbf{2005} & 100 & -6.177 & 11.12 & -2.059 & 93.86 \\
\textbf{2006} & 87.41 & -22.43 & 12.49 & -4.486 & 64.98 \\
\textbf{2007} & 52.56 & -18.87 & 6.57 & -6.289 & 33.7 \\
\textbf{2008} & 91.7 & -26.03 & 10.19 & -8.678 & 65.67 \\
\textbf{2009} & 57.69 & -40.85 & 11.54 & -5.836 & 16.84 \\
\textbf{2010} & 25.07 & -52.13 & 5.014 & -8.688 & -27.05 \\
\textbf{2011} & 6.812 & -21.58 & 6.812 & -10.79 & -14.77 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-8} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-cloudy-cond.}{%
\paragraph{Monthly cum sums for Cloudy cond.}\label{monthly-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1392}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1996} & 5.182 & -29.29 & 5.182 & -14.64 & -24.11 \\
\textbf{1997} & 37.41 & -8.749 & 18.7 & -4.375 & 28.66 \\
\textbf{1998} & 25.8 & -6.183 & 8.599 & -3.092 & 19.61 \\
\textbf{2000} & 9.461 & -9.806 & 9.461 & -9.806 & -0.3458 \\
\textbf{2001} & 19.95 & -9.823 & 19.95 & -9.823 & 10.13 \\
\textbf{2002} & 42.68 & -10.76 & 21.34 & -5.378 & 31.93 \\
\textbf{2003} & 35.13 & -21.3 & 17.57 & -10.65 & 13.83 \\
\textbf{2005} & 7.114 & -6.776 & 3.557 & -3.388 & 0.3376 \\
\textbf{2006} & 27.81 & -19.11 & 9.269 & -9.554 & 8.698 \\
\textbf{2008} & 5.68 & -15.25 & 2.84 & -5.085 & -9.574 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{cumulative-sums-by-sza-and-part-of-day}{%
\subsubsection{Cumulative sums by SZA and part of day}\label{cumulative-sums-by-sza-and-part-of-day}}

Use deseasonalized monthly values to calculate cumulative sums
by a SZA bin and a period of day.

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-3} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-6} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-9} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza-1}}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza-1}}

\textbf{END}

\begin{verbatim}
2023-06-25 21:53:08.2 athan@sagan DHI_GHI_1_longterm_trends.R 0.759222 mins
\end{verbatim}

\end{document}