### 2023-06-30 DHI_GHI_3_trends_consistency.tex
Commit 2023-06-30 13:17

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Trends of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-06-30}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI and DNI radiation for `clear sky' and all sly conditions.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
fANCOVA 0.6-1 loaded
\end{verbatim}

\begin{verbatim}
Loading required package: zoo
\end{verbatim}

\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}

Load environment and data from:  ~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda 
\end{verbatim}

\hypertarget{consistency-of-trends}{%
\subsection{3. Consistency of trends}\label{consistency-of-trends}}

\hypertarget{data-info}{%
\subsubsection{Data info}\label{data-info}}

Time data span 1993-04-12, 2023-03-31

Where is a \textbf{running mean the window is 1826.2125 days} or
5 years.

\hypertarget{process}{%
\subsubsection{Process}\label{process}}

\newpage
\FloatBarrier

\hypertarget{whole-day-daily-cumulative-sum}{%
\subsubsection{Whole day daily cumulative sum}\label{whole-day-daily-cumulative-sum}}

All data in this section are referring to daily means.

\newpage

\hypertarget{daily-cum-sums-for-all-sky-cond.}{%
\paragraph{Daily cum sums for All sky cond.}\label{daily-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 2212 & -3206 & 15.8 & -31.74 & -994.1 \\
\textbf{1994} & 4436 & -5582 & 22.75 & -35.33 & -1147 \\
\textbf{1995} & 3391 & -5423 & 21.06 & -30.81 & -2032 \\
\textbf{1996} & 2763 & -7748 & 16.55 & -41.65 & -4984 \\
\textbf{1997} & 4443 & -6187 & 22.78 & -38.91 & -1744 \\
\textbf{1998} & 4363 & -6104 & 22.04 & -40.16 & -1741 \\
\textbf{1999} & 4678 & -5052 & 22.38 & -33.02 & -374.3 \\
\textbf{2000} & 5675 & -3633 & 23.26 & -32.44 & 2042 \\
\textbf{2001} & 4554 & -5388 & 23 & -36.9 & -834.3 \\
\textbf{2002} & 4522 & -5948 & 23.55 & -36.05 & -1426 \\
\textbf{2003} & 4966 & -5994 & 24.34 & -38.67 & -1028 \\
\textbf{2004} & 4484 & -5385 & 21.56 & -36.39 & -900.9 \\
\textbf{2005} & 4594 & -3988 & 23.56 & -29.33 & 605.8 \\
\textbf{2006} & 4756 & -4391 & 23.54 & -33.02 & 364.7 \\
\textbf{2007} & 4437 & -4134 & 21.86 & -36.26 & 303.3 \\
\textbf{2008} & 5069 & -4850 & 21.94 & -38.19 & 218.9 \\
\textbf{2009} & 5239 & -5062 & 26.2 & -35.65 & 177.4 \\
\textbf{2010} & 3741 & -4367 & 23.09 & -36.39 & -625.9 \\
\textbf{2011} & 4773 & -4639 & 22.41 & -38.34 & 134.8 \\
\textbf{2012} & 4936 & -4286 & 21.46 & -40.43 & 650.9 \\
\textbf{2013} & 5311 & -4367 & 24.14 & -33.33 & 944.7 \\
\textbf{2014} & 4244 & -5710 & 21.44 & -38.06 & -1465 \\
\textbf{2015} & 5603 & -4512 & 27.07 & -36.09 & 1091 \\
\textbf{2016} & 6251 & -4629 & 27.42 & -34.54 & 1622 \\
\textbf{2017} & 6650 & -4326 & 26.92 & -37.94 & 2324 \\
\textbf{2018} & 5147 & -5090 & 24.51 & -34.86 & 57.66 \\
\textbf{2019} & 5454 & -3793 & 23.31 & -32.98 & 1661 \\
\textbf{2020} & 5971 & -3796 & 25.41 & -33.89 & 2175 \\
\textbf{2021} & 5726 & -4535 & 24.37 & -35.16 & 1191 \\
\textbf{2022} & 6210 & -4080 & 25.25 & -35.48 & 2131 \\
\textbf{2023} & 2356 & -752.9 & 36.24 & -35.85 & 1603 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-4} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-clear-sky-cond.}{%
\paragraph{Daily cum sums for Clear sky cond.}\label{daily-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 2792 & -3585 & 21.32 & -33.2 & -793 \\
\textbf{1994} & 5678 & -5941 & 28.68 & -38.83 & -263.5 \\
\textbf{1995} & 3992 & -6119 & 28.11 & -31.54 & -2127 \\
\textbf{1996} & 3784 & -8260 & 25.23 & -40.69 & -4476 \\
\textbf{1997} & 4743 & -6824 & 26.95 & -38.34 & -2081 \\
\textbf{1998} & 4676 & -6602 & 25.83 & -39.06 & -1926 \\
\textbf{1999} & 5571 & -5569 & 28.14 & -33.96 & 1.884 \\
\textbf{2000} & 5944 & -4083 & 26.54 & -30.93 & 1862 \\
\textbf{2001} & 4425 & -5888 & 24.72 & -36.57 & -1463 \\
\textbf{2002} & 5191 & -6541 & 30.18 & -35.74 & -1350 \\
\textbf{2003} & 5324 & -6574 & 27.87 & -39.13 & -1250 \\
\textbf{2004} & 4656 & -5972 & 26.45 & -33.18 & -1316 \\
\textbf{2005} & 5603 & -4494 & 30.95 & -30.16 & 1109 \\
\textbf{2006} & 5674 & -4808 & 30.34 & -32.71 & 866.5 \\
\textbf{2007} & 5235 & -4432 & 27.26 & -35.46 & 802.4 \\
\textbf{2008} & 5972 & -5227 & 28.3 & -35.8 & 745 \\
\textbf{2009} & 5518 & -5543 & 29.99 & -35.08 & -24.72 \\
\textbf{2010} & 3571 & -4758 & 24.63 & -34.73 & -1187 \\
\textbf{2011} & 4798 & -5062 & 24.99 & -35.9 & -264.2 \\
\textbf{2012} & 5330 & -4685 & 25.63 & -36.89 & 645.1 \\
\textbf{2013} & 5214 & -4820 & 25.43 & -33.01 & 394.5 \\
\textbf{2014} & 4786 & -6220 & 25.87 & -38.16 & -1433 \\
\textbf{2015} & 5689 & -4928 & 29.17 & -35.97 & 761.3 \\
\textbf{2016} & 6793 & -5050 & 30.74 & -36.07 & 1743 \\
\textbf{2017} & 6491 & -4761 & 28.98 & -34.75 & 1730 \\
\textbf{2018} & 5912 & -5647 & 30.63 & -34.65 & 264.4 \\
\textbf{2019} & 5854 & -4293 & 28.01 & -30.88 & 1561 \\
\textbf{2020} & 6068 & -4308 & 28.76 & -31.68 & 1759 \\
\textbf{2021} & 6665 & -4972 & 29.75 & -35.51 & 1693 \\
\textbf{2022} & 6786 & -4448 & 28.75 & -35.58 & 2338 \\
\textbf{2023} & 2484 & -806.3 & 40.07 & -35.06 & 1678 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-8} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-cloudy-cond.}{%
\paragraph{Daily cum sums for Cloudy cond.}\label{daily-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 555.1 & -1190 & 20.56 & -27.05 & -635.2 \\
\textbf{1994} & 661.9 & -1756 & 18.39 & -24.73 & -1094 \\
\textbf{1995} & 633.6 & -1041 & 18.1 & -21.68 & -407.2 \\
\textbf{1996} & 873 & -1720 & 26.46 & -28.2 & -847.3 \\
\textbf{1997} & 1256 & -1106 & 23.25 & -26.97 & 149.8 \\
\textbf{1998} & 1047 & -1000 & 20.53 & -20.84 & 46.84 \\
\textbf{1999} & 713.1 & -1381 & 17.39 & -20.93 & -668.3 \\
\textbf{2000} & 964.4 & -1159 & 22.96 & -22.72 & -194.5 \\
\textbf{2001} & 1006 & -655.4 & 24.54 & -18.73 & 350.7 \\
\textbf{2002} & 1168 & -1153 & 21.23 & -23.52 & 15.3 \\
\textbf{2003} & 1044 & -1142 & 23.74 & -23.31 & -97.64 \\
\textbf{2004} & 1356 & -645.1 & 27.67 & -19.55 & 711 \\
\textbf{2005} & 825.9 & -1101 & 20.14 & -20.78 & -275.5 \\
\textbf{2006} & 941.5 & -1016 & 21.9 & -22.08 & -74.35 \\
\textbf{2007} & 655.5 & -1183 & 21.85 & -25.72 & -527.7 \\
\textbf{2008} & 844.9 & -1397 & 22.83 & -21.82 & -551.9 \\
\textbf{2009} & 830.6 & -752.8 & 22.45 & -22.14 & 77.77 \\
\textbf{2010} & 792.5 & -529.8 & 23.31 & -21.19 & 262.7 \\
\textbf{2011} & 1309 & -776.5 & 26.19 & -23.53 & 532.8 \\
\textbf{2012} & 1135 & -1052 & 25.22 & -25.04 & 83.18 \\
\textbf{2013} & 1382 & -776 & 27.1 & -18.05 & 606.3 \\
\textbf{2014} & 856.1 & -819.9 & 23.14 & -20 & 36.18 \\
\textbf{2015} & 948.2 & -515.4 & 21.07 & -20.61 & 432.8 \\
\textbf{2016} & 1287 & -1040 & 24.74 & -22.13 & 246.5 \\
\textbf{2017} & 972.8 & -737.2 & 22.62 & -21.06 & 235.6 \\
\textbf{2018} & 1244 & -871.7 & 24.4 & -23.56 & 372.6 \\
\textbf{2019} & 921.7 & -863.8 & 23.63 & -21.07 & 57.86 \\
\textbf{2020} & 1194 & -583.9 & 22.12 & -17.69 & 610.5 \\
\textbf{2021} & 1043 & -844.4 & 16.55 & -16.56 & 198.3 \\
\textbf{2022} & 1113 & -784 & 21.83 & -19.12 & 329.3 \\
\textbf{2023} & 105.6 & -88.26 & 13.21 & -22.07 & 17.38 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{whole-day-monthly-cumulative-sum}{%
\subsubsection{Whole day monthly cumulative sum}\label{whole-day-monthly-cumulative-sum}}

In this section the monthly values have been calculated from daily mean values.

\newpage

\hypertarget{monthly-cum-sums-for-all-sky-cond.}{%
\paragraph{Monthly cum sums for All sky cond.}\label{monthly-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 19.08 & -49.18 & 6.36 & -9.835 & -30.1 \\
\textbf{1994} & 25.68 & -74.79 & 6.42 & -9.348 & -49.11 \\
\textbf{1995} & 26.96 & -93.72 & 5.391 & -15.62 & -66.76 \\
\textbf{1996} & 3.185 & -182.8 & 1.592 & -18.28 & -179.7 \\
\textbf{1997} & 25.69 & -81.23 & 8.565 & -9.025 & -55.53 \\
\textbf{1998} & 27.98 & -76.69 & 5.596 & -10.96 & -48.71 \\
\textbf{1999} & 7.85 & -24.89 & 1.962 & -3.111 & -17.04 \\
\textbf{2000} & 77 & -18.23 & 8.556 & -6.078 & 58.77 \\
\textbf{2001} & 49.51 & -77.16 & 9.901 & -11.02 & -27.66 \\
\textbf{2002} & 38.91 & -85.37 & 12.97 & -9.486 & -46.46 \\
\textbf{2003} & 29.62 & -56.58 & 5.924 & -8.083 & -26.97 \\
\textbf{2004} & 28.07 & -57.25 & 4.679 & -9.541 & -29.17 \\
\textbf{2005} & 33.92 & -15.39 & 4.845 & -3.847 & 18.53 \\
\textbf{2006} & 36.39 & -35.02 & 7.278 & -5.836 & 1.371 \\
\textbf{2007} & 24.11 & -15.83 & 4.821 & -3.165 & 8.282 \\
\textbf{2008} & 49.93 & -31.83 & 7.133 & -6.367 & 18.09 \\
\textbf{2009} & 55.82 & -55.67 & 7.974 & -11.13 & 0.1516 \\
\textbf{2010} & 24.08 & -38.35 & 6.02 & -9.587 & -14.27 \\
\textbf{2011} & 24.81 & -29.36 & 4.134 & -5.872 & -4.555 \\
\textbf{2012} & 44.97 & -24.08 & 5.622 & -8.025 & 20.9 \\
\textbf{2013} & 64.22 & -36.75 & 8.027 & -9.188 & 27.47 \\
\textbf{2014} & 19.91 & -64.78 & 3.982 & -9.255 & -44.87 \\
\textbf{2015} & 92.08 & -45.23 & 13.15 & -11.31 & 46.86 \\
\textbf{2016} & 72.74 & -18.02 & 8.083 & -6.007 & 54.72 \\
\textbf{2017} & 89.13 & -12.27 & 11.14 & -3.067 & 76.87 \\
\textbf{2018} & 50.91 & -44.67 & 8.484 & -7.446 & 6.233 \\
\textbf{2019} & 61.05 & -17 & 8.721 & -3.399 & 44.05 \\
\textbf{2020} & 96.81 & -20.9 & 10.76 & -6.966 & 75.91 \\
\textbf{2021} & 60.17 & -21.36 & 7.521 & -5.34 & 38.81 \\
\textbf{2022} & 93.42 & -18.67 & 11.68 & -4.668 & 74.75 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-4} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-clear-sky-cond.}{%
\paragraph{Monthly cum sums for Clear sky cond.}\label{monthly-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 23 & -44.04 & 7.667 & -8.808 & -21.04 \\
\textbf{1994} & 43.71 & -62 & 7.286 & -10.33 & -18.29 \\
\textbf{1995} & 32.7 & -101.9 & 5.449 & -16.98 & -69.17 \\
\textbf{1996} & 12.59 & -175.8 & 2.518 & -25.11 & -163.2 \\
\textbf{1997} & 24.27 & -91.54 & 8.089 & -10.17 & -67.27 \\
\textbf{1998} & 26.02 & -81.07 & 5.204 & -11.58 & -55.05 \\
\textbf{1999} & 17.75 & -23.97 & 4.437 & -2.997 & -6.225 \\
\textbf{2000} & 70.16 & -18.5 & 8.77 & -4.626 & 51.66 \\
\textbf{2001} & 43.97 & -96.69 & 7.328 & -16.12 & -52.72 \\
\textbf{2002} & 39.55 & -84.05 & 13.18 & -9.339 & -44.5 \\
\textbf{2003} & 23.56 & -58.22 & 7.854 & -6.469 & -34.66 \\
\textbf{2004} & 23.1 & -65.25 & 7.7 & -7.25 & -42.15 \\
\textbf{2005} & 72.04 & -8.362 & 8.004 & -2.787 & 63.68 \\
\textbf{2006} & 65.65 & -30.27 & 9.379 & -6.053 & 35.39 \\
\textbf{2007} & 38.64 & -25.59 & 5.52 & -6.397 & 13.05 \\
\textbf{2008} & 69.2 & -32.11 & 8.65 & -8.028 & 37.09 \\
\textbf{2009} & 48.84 & -58.31 & 8.14 & -9.718 & -9.469 \\
\textbf{2010} & 17.62 & -63.38 & 3.523 & -10.56 & -45.76 \\
\textbf{2011} & 21.77 & -42.12 & 5.443 & -6.017 & -20.35 \\
\textbf{2012} & 44.72 & -23.67 & 6.388 & -5.917 & 21.05 \\
\textbf{2013} & 57.4 & -48.51 & 8.2 & -9.703 & 8.89 \\
\textbf{2014} & 16.72 & -60.33 & 3.344 & -8.618 & -43.61 \\
\textbf{2015} & 87.63 & -47.43 & 14.6 & -7.905 & 40.2 \\
\textbf{2016} & 86.79 & -24.92 & 8.679 & -12.46 & 61.87 \\
\textbf{2017} & 80.29 & -22.9 & 11.47 & -4.58 & 57.38 \\
\textbf{2018} & 59.82 & -47.65 & 9.97 & -7.942 & 12.17 \\
\textbf{2019} & 60.07 & -20.82 & 10.01 & -3.47 & 39.26 \\
\textbf{2020} & 93.94 & -31.97 & 15.66 & -5.329 & 61.97 \\
\textbf{2021} & 78 & -21.81 & 9.75 & -5.452 & 56.19 \\
\textbf{2022} & 105.1 & -22.41 & 13.14 & -5.602 & 82.71 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-8} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-cloudy-cond.}{%
\paragraph{Monthly cum sums for Cloudy cond.}\label{monthly-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1996} & 1.041 & -32.43 & 1.041 & -16.21 & -31.39 \\
\textbf{1997} & 30.94 & -13.41 & 15.47 & -6.707 & 17.52 \\
\textbf{1998} & 22.07 & -11.6 & 7.356 & -5.802 & 10.46 \\
\textbf{2000} & 5.152 & -11.35 & 5.152 & -11.35 & -6.197 \\
\textbf{2001} & 17.63 & -11.37 & 17.63 & -11.37 & 6.265 \\
\textbf{2002} & 35.68 & -15.21 & 17.84 & -7.604 & 20.47 \\
\textbf{2003} & 31.25 & -26.37 & 15.62 & -13.19 & 4.878 \\
\textbf{2005} & 0.5237 & -13.4 & 0.5237 & -4.466 & -12.87 \\
\textbf{2006} & 20.46 & -22.56 & 6.82 & -11.28 & -2.099 \\
\textbf{2008} & 0.8906 & -22.67 & 0.4453 & -7.558 & -21.78 \\
\textbf{2011} & 9.071 & -7.749 & 4.535 & -7.749 & 1.322 \\
\textbf{2012} & 18.89 & -26.75 & 9.443 & -26.75 & -7.868 \\
\textbf{2013} & 23.68 & -3.169 & 11.84 & -3.169 & 20.51 \\
\textbf{2016} & 15.11 & -12.11 & 7.554 & -4.037 & 2.997 \\
\textbf{2021} & 15.2 & -16.57 & 5.068 & -8.284 & -1.366 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{cumulative-sums-by-sza-and-part-of-day}{%
\subsubsection{Cumulative sums by SZA and part of day}\label{cumulative-sums-by-sza-and-part-of-day}}

Use deseasonalized monthly values to calculate cumulative sums
by a SZA bin and a period of day.

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-3} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-6} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-9} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-6} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-12} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-18} \end{center}

\textbf{END}

\begin{verbatim}
2023-06-30 09:47:23.0 athan@sagan DHI_GHI_1_longterm_trends.R 1.903256 mins
\end{verbatim}

\end{document}