### 2023-07-03 Article.tex
Commit 2023-07-03 12:17

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Trends of SDR in Thessaloniki},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{subcaption}

\begin{document}

\begin{frontmatter}

  \title{Trends of SDR in Thessaloniki}
    \author[Laboratory of Atmospheric Physics]{Athanasios Natsis%
  \corref{cor1}%
  }
   \ead{natsisphysicist@gmail.com} 
    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
      \affiliation[Laboratory of Atmospheric Physics]{
    organization={Aristotle University of
Thessaloniki},addressline={Campus Box
149},city={Thessaloniki},postcode={54124},country={Greece},}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Study of GHI and DNI radiation for `clear-sky' and all-sky conditions.
  It consists of two paragraphs.
  \end{abstract}
    \begin{keyword}
    GHI \sep SDR \sep 
    Solar Brigthening/Dimming
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction.}{%
\section{Introduction.}\label{introduction.}}

The shortwave downward solar irradiance (SDR) at Earth's surface play a
significant role, on it's climate. Changes of the SDR can be related to
changes on Earth's energy budget, the mechanisms of climate change, and
water and carbon cycle \citep{Wild2009}. Can also affect, solar and
agricultural production, and living organisms. Studies of SDR
variability, have identified some distinct SDR trends on different
regions of the world and for different time periods. The term
`brightening' is generally used to describe periods of positive SDR
trend, and `dimming' for negative trend. There are many cases on the
long term records of irradiance, showing a systematic change of SDR's
trend slope, occurring roughly at the last decades of the 20th century.
On multiple station in China, a dimming period was reported until about
2000, followed by a brightening period \citep{Yang2021}. A similar
pattern was identified, with the breaking point around 1980, for
stations in Central Europe \citep{Wild2021} and Brazil
\citep{Yamasoe2021}. Also, on global scale, an AI aided continental
level spatial analysis, with data from multiple station, reach similar
conclusions for the above regions and for the global trend
\citep{Yuan2021}.

There is a consensus, among researchers, that major factors of SDR
attenuation is the interaction with atmospheric aerosols and clouds.
Those interactions, among other factors, have been analysed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar finds, have been showed in observational data
\citep[ and references
therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.

Due to the variability of the phenomenon, and its contributing factors,
there is a need to further investigate and monitor SDR, in different
sites, to estimate its magnitude, and its relation to the local
conditions. In this study, we examine the trends of SDR, with
ground-based measurements at Thessaloniki, Greece for the period 1993 to
2023, as derived from a CM-21 pyranometer. We reevaluated and extended
the dataset used by \citet{Bais2013}, applying a different algorithm for
the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
period 1993 to 2023, under different sky conditions (all-sky, clear-sky
and cloud-sky).

\hypertarget{observational-data-and-methodology.}{%
\section{Observational data and
methodology.}\label{observational-data-and-methodology.}}

The SDR is equivalent to the whole sky Global radiation (or Global
Horizontal Irradiance, GHI), and was obtained with a horizontal leveled,
CM-21 pyranometer, from 1993-04-12 to 2023-03-31. Also, we have
measurements of direct beam radiation (DNI) obtained by a collocated
CHP-1 pyrheliometer, from 2016-04-01 to 2023-03-31. Although, we
performed a similar analysis to the DNI as for the SDR, the results are
not presented here, as they lack the appropriate statistical
significance, due to the sorter duration of the data. Moreover, the DNI
was used as auxiliary data, with the clear sky identification algorithm
(CSid), to aid with the selection of the appropriate thresholds. We have
to note also, that despite the capability of the CSid to use the DNI as
an characterization parameter, we haven't use it, to avoid any selection
bias due to unequal availability of measurements.

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ38'\)N,
\(22^\circ57'\)E, 80m a.s.l.). The monitoring site is located near the
city centre. During the study period the CM-21 pyranometer has been
independently calibrated four times. There are three distinct steps to
the creation of this dataset: a) the acquisition of radiation
measurements from the sensors, b) a radiation data quality check, and c)
the identification of ``clear sky'' conditions from the radiometric data
d) data aggregation and trend analysis.

For the acquisition of radiometric data, the signal of the broadband
instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its
standard deviation are recorded for every minute. The measurements are
corrected for the zero offset of the instrument signal. As reference of
the ``dark signal'' is used the Sun elevation angle bellow
\(-10^\circ\), for a period of \(3 \text{h}\). The signal is converted
to radiation flux, using a ramped value of the instrument sensitivity,
derived from the eight laboratory calibrations of the instrument, during
the study period.

A manual screening was performed, to remove inconsistencies and
erroneous recordings, that can occur randomly or systematically, during
the long continuous operation of the instrument. The manual screening is
aided by a radiation data quality assurance procedure, adjusted for this
site, based on methods of Long and
Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
problematic recordings have been excluded from further processing.
Although it is not possible to detect all the bad data, the large number
of data and the aggregation scheme we use, can provide us with accurate
radiometric measurements, for the scope of this study.

In order to estimate the effect of the clouds on the SDR, we created
three datasets, by characterizing each one-minute measurement with a
corresponding sky condition. The all-sky conditions, containing all the
valid measurements. The clear-sky conditions data, correspond to data
where we have inferred that the sky was almost clear of clouds, and the
remainder part the data was characterized with cloud sky conditions
(cloud-sky). To identify the clear-sky conditions we used the method
proposed by \citet{Long2000} and by \citet{Reno2016}, that was adapted
and configured for the site.

\ldots.

cite iterative process

\ldots\ldots{}

We have to note, that the definition of what constitutes as clear or
cloudy sky, has some subjectivity, due to the used method of
characterization. As a result, the details of the definition are site
specific, it relies on a combination of threshold of comparisons with
ideal actinometric models and statistics on different signal behavior.

The algorithm was calibrated with the main focus to identify the
presence of clouds on the sky dome, although there are marginal cases
that there will be false positives or false negatives identifications.

As global radiation clear sky reference we are using the Haurwitz's
model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
The selection of a clear sky reference model, was based on SDR
observation from the period 2016 -- 2021. Where, after an iterative
optimization of eight simple models (Daneshyar--Paltridge--Proctor,
Kasten--Czeplak, Haurwitz, Berger--Duffie,
Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and
Ineichen-Perez) with different factors. We found, that Haurwitz's model,
adjusted with a factor of 0.965 has the lower root mean squared error
(RMSE). The tried models are described by \citet{Reno2012} and tested by
\citet{Reno2016}. The iterative optimization method, for the selection
of the reference is discussed by \citet{Long2000} and \citet{Reno2016}.

\begin{equation}
\text{SDR}_\text{Clear Sky} = 0.965 \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
\end{equation}

The following criteria and thresholds were used to identify clear-sky
conditions. Each criterion was applied for a running window of \(11\)
consecutive measurements/minutes, and the characterization is applied at
the center value of the window. A data point is consider as under
cloud-sky condition if it fails to pass any of the criteria, all the
other data points are characterized as clear-sky.

The mean of the measured value \(\overline{G}_i\) must be inside an
envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\)
(Eq. \ref{eq:MeanVIP}).

\begin{equation}
0.91 \times \overline{\text{SDR}}_{i\text{Clear Sky}} - 20
< \overline{G}_i <
1.095 \times \overline{\text{SDR}}_{i\text{Clear Sky}} + 30
\label{eq:MeanVIP}
\end{equation}

The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
compared to a similar constructed value from the reference
\(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).

\begin{equation}
1 \times M_{CSi} - 75
< M_{Gi} <
1 \times M_{CSi} + 75
\label{eq:MaxVIP}
\end{equation}

The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments
connecting the points of the SDR time series for the measured values
\(L\) and similar for the reference \(L_{CS}\), must be within the
limits of Eq. \ref{eq:VILcr}.

\begin{equation}
L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}

\begin{equation}
1 \times L_{CSi} - 5 < L_i < 1.3 \times L_{CSi} + 13
\label{eq:VILcr}
\end{equation}

We calculate the standard deviation \(\sigma\) of the slope (\(s\))
between sequential points in the time series, normalized by the average
SDR during the time interval.

\begin{equation}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
\end{equation}

\begin{equation}
\sigma_i < \ensuremath{1.1\times 10^{-4}}
\label{eq:VCTcr}
\end{equation}

The maximum difference \(X\) between the change in measured irradiance
and the change in clear sky irradiance over each measurement interval.

\begin{gather}
x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
\end{gather}

\begin{equation}
X_i < 7.5
\label{eq:VSMcr}
\end{equation}

Due to a significant measurement uncertainty near, the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all 1-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
the Sun's intensity variability, using a homogenized time series of
satellite TSI observations. The TSI data we use, comes from NOAA
\citep{Coddington2005}, it overlaps with our data, and is interpolated
to match with the time step of the data. As a result, we can present all
radiation data as a fraction of Sun's TSI.

\hypertarget{aggregation-of-radiometric-data}{%
\subsubsection{Aggregation of radiometric
data}\label{aggregation-of-radiometric-data}}

Before further analysis and deseasonalization, we implement an
appropriate aggregation scheme on the 1-minute data. To preserve the
representativeness we use the following criteria: a) for daily mean
values we exclude instances with less than 180 valid data points, b)
accordingly, monthly values are computed by daily aggregated data, where
months with less than 20 days are rejected. The daily seasonal values
were indexed based on the day of year number, and the monthly by the
corresponding calendar month. For the seasons of the year trends, we
grouped the mean daily values by season (December - February: Winter,
March - May: Spring, etc.). Finally, for each data set, we remove the
natural occurring seasonal variation. This is done, by calculating the
mean values for the appropriate time step, and subtracting the annual
cycle from the actual data. In addition to the previous aggregation
scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for
cases before and after local noon. This gave us an approximation of
Sun's SZA contribution to the SDR brightening.

Thus, for each data set we have obtained the relative departure of the
seasonal mean. The statistical significance and aggregation scheme and
filtering will be noted along with the corresponding results.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{long-term-sdr-trends}{%
\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}

We calculated the SDR trends, from the mean daily values of the
departure from the seasonal values, for the whole time series for the
three sky conditions (Table \ref{tab:trendtable}).

\begin{figure}[h!]

{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}

}

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
\end{figure}

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
\caption{\label{tab:trendtable}Trends of daily SDR means, for different
sky conditions. All cases have a statistical significance above
\(99.9\%\) \normalsize}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.3996 & 8.665e-23 & All sky cond. \\
0.3773 & 7.075e-18 & Clear sky cond. \\
0.3594 & 3.318e-09 & Cloudy cond. \\
\end{longtable}

\hypertarget{long-term-trends-by-solar-zenith-angle.}{%
\subsection{Long term trends by Solar zenith
angle.}\label{long-term-trends-by-solar-zenith-angle.}}

Analyzing the long term trends of SDR to bins of SZA, we can see the
contribution of the geometry and time in a diurnal level (Figures
\ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a
seasonal distribution of SZA than in not presented here.

\begin{figure}[h!]

{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

}

\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-trends-by-season-of-year}{%
\subsection{Long term trends by season of
year}\label{long-term-trends-by-season-of-year}}

Similar we have produced the deseasonalized trend for different sky
conditions for each season of the year, using the corresponding mean
monthly values. This can give us a better understanding of the annual
variability of the trends.

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether-1}

}

\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
\caption{\label{tab:trendseasontable}Trends of daily means by sky
conditions for the seasons of the year. (continued
below)}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.7215 & 0.001209 & All sky cond. & Winter \\
0.1537 & 0.1498 & All sky cond. & Spring \\
0.128 & 0.08705 & All sky cond. & Summer \\
0.4239 & 0.006659 & All sky cond. & Autumn \\
0.8715 & 0.000269 & Clear sky cond. & Winter \\
0.05881 & 0.6247 & Clear sky cond. & Spring \\
-0.02369 & 0.7972 & Clear sky cond. & Summer \\
0.3583 & 0.02996 & Clear sky cond. & Autumn \\
-0.02581 & 0.9335 & Cloudy cond. & Winter \\
0.4333 & 0.02937 & Cloudy cond. & Spring \\
0.442 & 0.004988 & Cloudy cond. & Summer \\
0.4776 & 0.01955 & Cloudy cond. & Autumn \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
99.88 \\
85.02 \\
91.29 \\
99.33 \\
99.97 \\
37.53 \\
20.28 \\
97 \\
6.646 \\
97.06 \\
99.5 \\
98.04 \\
\end{longtable}

\hypertarget{consistency-of-the-trends}{%
\subsection{Consistency of the trends}\label{consistency-of-the-trends}}

A method to evaluate changes in the long term trend is to use the
cumulative sum of the variable.

Figure \ref{fig:cumsum}

\ldots{} different in aggregation \ldots{}

\begin{figure}[h!]

{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

}

\caption{two plots}\label{fig:cumsum}
\end{figure}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous
results of \citet{Bais2013} for the site. The increase of this trend,
shows that the phenomenon and probably the causes have been amplified.

\ldots.. Also, the trend of \(0.36\%/year\) for cloud-sky condition
indicate the major part that clouds play on the \ldots..

\ldots\ldots.

About:

\begin{itemize}
\item
  long terms trends
\item
  seasonal trends
\item
  effect of SZA
\item
  effect of clouds
\item
  Aerosols
\end{itemize}

similar results with Evidence for Clear‐Sky Dimming and Brightening in
Central Europe\_Wild2021.pdf inclue in discusion

\begin{longtable}[]{@{}c@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{END} \\
\end{longtable}

\bibliography{Mainreferences.bib}

\end{document}