### 2023-07-13 Build_151/Article_B151.Rmd
Commit 2023-07-13 08:09

---
title:         "Trends of SDR in Thessaloniki"
# header-includes:
author:
  - name:         Athanasios Natsis 
    email:        natsisphysicist@gmail.com
    affiliation:  Laboratory of Atmospheric Physics
#    footnote:     1
    correspondingauthor: true
  - name:         Alkiviadis Bais
    email:        abais@auth.gr 
    affiliation:  Laboratory of Atmospheric Physics
    correspondingauthor: false
address:
  - code:         Laboratory of Atmospheric Physics
    organization: Aristotle University of Thessaloniki
    addressline:  Campus Box 149
    city:         Thessaloniki
    # state:        State
    postcode:     54124
    country:      Greece
#  - code:         Another University
#    organization: Department
#    addressline:  A street 29
#    postcode:     2054 NX
#    city:         Manchester,
#    country:      The Netherlands
# footnote:
#   - code:         1
#     text:         "This is the first author footnote."
#   - code:         2
#     text:         "Another author footnote."
abstract: |
  Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.
  It consists of two paragraphs.
keywords: 
  - GHI
  - SDR
  - Solar Brigthening/Dimming
journal:         "An awesome journal"
date:            "`r Sys.Date()`"
classoption:     preprint, 3p, authoryear
bibliography:    Mainreferences.bib
linenumbers:     false
numbersections:  true
header-includes:
   # - \usepackage{subfig}
   - \usepackage{subcaption}

output: 
  rticles::elsevier_article:
    keep_tex:         no
    citation_package: natbib
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              no
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
  bookdown::word_document2: default
# date:           "`r format(Sys.time(), '%F')`"
---

```{r echo=F, include=T, warning=FALSE, message=FALSE}
suppressMessages(
    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
)
## document configurations 
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "70%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
## load data
load("./data/common_data_14_2.Rda")
## number of measurements
SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])

TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
                    strip.white = TRUE,
                    sep = ";",
                    header = TRUE))
## variables we use
source("./DHI_GHI_0_variables.R")

## create local citations file and links here
source("~/CODE/FUNCTIONS/R/gather_citations.R")

suppressMessages(
    gather_citations(
        sourcefile   = "./Article.Rmd",
        mainreposit  = "./Mainreferences.bib",
        localreposit = "./references.bib",
        folderlink   = "./References_links",
        workingdir   = "./"
    )
)
```

<!-- use SDR for GHI -->

# Introduction.

<!-- something about aerosols: bais mail -->
<!-- something about clouds: bais mail -->
<!--
Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
-->
....

more about trend in relation to other trends
Aerosols and clouds

......

The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Can also affect, solar and agricultural production, and living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
<!-- Trends china -->
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
<!-- Trends Europe Brazil -->
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
<!-- Trends Global -->
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].

There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds. 
Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->

Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
<!-- goals -->
We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).

<!--
Aerosols interactions

#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China

Aerosols not only strengthen but also weaken the
growth of clouds with respect to their coverage and
optical thickness, depending on the levels of pollution
and the associated amounts of aerosols as postulated
in a conceptual framework (Wild 2009a, 2012, Yang
et al 2012).

...... cloud formation occurrences and constitution nuclei. cite .....
p>        "  Too hard to vim

#Wild2012 : simplify interaction nice graph

-->

# Observational data and methodology.

<!-- time span -->
The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, in the period from
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS$Date)`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS$Date)`.
<!-- }. -->
Along with SDR, we obtained the direct beam radiation (DNI) by a collocated CHP-1 pyrheliometer, for the period
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
<!-- }. -->
Although, we have performed a similar analysis to the DNI, as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
Moreover, the DNI data were used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
We have to note, that despite the capability of the CSid algorithm, to use the DNI as a characterization parameter, we haven't used it here, to avoid any selection bias due to unequal availability of the two datasets.

<!-- Location -->

The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.).
The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
During the study period, the CM-21 pyranometer has been independently calibrated ..... **four** .... times.
........ verify number ........
<!-- Preparation. -->
There are three distinct steps in the creation of the dataset analysed here:
a)\ the acquisition of radiation measurements from the sensors, 
b)\ a radiation data quality check,
c)\ the identification of "clear sky" conditions from the radiometric data, and
d)\ data aggregation and trend analysis.

<!--
The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.  These data points were excluded from the analysis, as
-->

<!-- Zerefos

Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
loniki, the major city in northern Greece, lies along the exit of air
pollutants from Europe, in the path of the prevailing long-range
transport (LRT) trajectories, as we move from central and east-
ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
-->

<!-- a) Acquisition -->
For the acquisition of radiometric data,
the signal of the broadband instruments is sampled with a rate of $1\,\text{Hz}$. The mean and the standard deviation are recorded for every minute.
<!-- dark correction -->
The measurements are corrected for the zero offset of the instrument dark voltage signal.
As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3\,\text{h}$.
<!-- physical quantities -->
The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the ...... **eight** ......  laboratory calibrations of the instrument, during the study period.

<!-- b) Radiation data quality check -->
<!-- The radiation quality is assured with a manual screening, aided by an automatic assessment process. -->
<!-- screening -->
A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur stochastically or systematically, during the continuous operation of the instrument.
<!-- Quality assurance -->
The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, based on methods of Long and Shi\ [-@Long2008a; -@Long2006].
Thus, problematic recordings have been excluded from further processing. 
Although it is not possible to detect all the bad data, the large number of available data, and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.

<!-- c) Clear sky identification -->
In order to be able to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
The all-sky conditions, containing all the valid measurements. 
The clear-sky conditions data, corresponding to data, where we have inferred that the sky was almost clear of clouds, and the remainder part of the data, characterized as cloud-sky conditions. 
To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site, as the authors suggest.

<!-- CS subjectivity -->
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in this method of characterization.
As a result, the details of the definition is site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
<!-- basic concepts -->
The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although few marginal cases exist that will be false positive or false negative identified, but can not affect the general outcome.

```{r setup, echo=F, include=T, eval=TRUE}
## Load Clear Sky id v14.2 variables
MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
# export all Clear Sky variables
for (an in names(MS)) {
assign(an, MS[[an]])
}
## get TSI slope for reference
TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
           strip.white = TRUE,
           sep = ";",
           header = TRUE))
TSI <- TSI[grep("TSI", TSI$var), ]
TSI$slope.sd           <- NULL
TSI$slope.ConfInt_0.95 <- NULL
TSI$slope.ConfInt_0.99 <- NULL
TSI$var                <- NULL
setorder(TSI, -slope.stat_sig)
rownames(TSI) <- NULL
names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
TSIskopeAll <- TSI[`Sky condition` == "All sky cond."]
```

<!-- # Detection of clear periods in GHI measurements for SDR trends. -->

For completeness, we will provide a brief overview of the CSid process, along with the site specific thresholds.
As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with a factor $a$, resulting by an iterative optimization process, as described by @Long2000 and @Reno2016.
The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], based on work by @Brent1973.
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 – 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors. We found, that Haurwitz’s model, adjusted with the factor $a = `r signif(alpha,3)`$ has one of the lowest root mean squared error (RMSE), while correlates with large amount of our data. Thus, our clear sky reference is the Eq.\ \ref{eq:ahau}. 
The tried models are described in more details by @Reno2012 and evaluated by @Reno2016.

<!-- cost fuction -->
\begin{equation}
f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i})^2 \label{eq:minf}
\end{equation}

<!-- adjusted Haurwitz -->
\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz}  = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
\end{equation}

The following criteria and thresholds were used to identify clear-sky conditions.
Each criterion was applied for a running window of $`r nt`$ consecutive measurements/minutes, and the characterization is applied at the center value of the window.
A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, the rest data points are characterized as clear-sky.

<!-- 1. MeanVIP -->
<!-- ### Mean value of irradiance during the time period. -->

The mean of the measured value $\overline{\text{SDR}}_i = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ must be inside an envelope based on the equivalent mean of the reference model $\overline{\text{SDR}}_\text{CSref} = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq. \ref{eq:MeanVIP}).

\begin{equation}
`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
< \overline{SDR}_i <
`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
\label{eq:MeanVIP}
\end{equation}

<!-- 2. MaxVIP -->
<!-- ### Max value of irradiance during the time period. -->

The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).

\begin{equation}
`r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
< M_{\text{SDR},i} <
`r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
\label{eq:MaxVIP}
\end{equation}

<!-- 3. VIL -->
<!-- ### Variability in irradiance by line length. -->

The length $L$ of the sequence of line segments, connecting the points of the SDR time series for the measured values (Eq.\ \ref{eq:VILeq}), and similar for the reference $L_{CS}$, must be within the limits of Eq. \ref{eq:VILcr}.

\begin{equation}
L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}

\begin{equation}
`r MinVIL_fct` \times L_{CSi} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{CSi} + `r offVIL_upl`
\label{eq:VILcr}
\end{equation}

<!-- 4. VCT -->
<!-- ### Variance of Changes in the Time series. -->

<!-- Standard deviation of rate of change in irradiance. -->
The standard deviation $\sigma$ of the slope ($s$) between sequential points in the time series, normalized by the average SDR during the time interval, must be bellow a threshold (Eq.\ \ref{eq:VCTcr}).

<!--
\begin{gather}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather}
-->

\begin{equation}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
\end{equation}

\begin{equation}
\sigma_i < `r offVCT`
\label{eq:VCTcr}
\end{equation}

<!-- 5. VSM -->
<!-- ### Variability in the Shape of the irradiance Measurements. -->

The maximum difference $X$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval, must be lower than a threshold (Eq.\ \ref{eq:VSMcr}).

\begin{gather}
x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3}
\end{gather}

\begin{equation}
X_i < `r offVSM`
\label{eq:VSMcr}
\end{equation}

<!-- ### Data selection -->
Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
On the latter instances, Sun is systematically, not visible by the instrument's location.
<!-- 1au -->
To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
<!-- TSI -->
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
The TSI data we use, comes from NOAA [@Coddington2005], it overlaps with our data, and is interpolated to match with the time step of the data.
<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
As a result, we can present all radiation data as a fraction of Sun's TSI. 
After the preparation of the data, we have $`r SDR_N_All`$  1-minute measurements, of which,
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
are characterized with clear-sky condition, and the rest 
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
with cloud-sky conditions.

<!-- aggregation -->
In order to investigate the SDR trend, we implement an appropriate aggregation scheme of the 1-minute data to a coarser time scale. To preserve the representativeness we are using the following criteria:
a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points,
b) accordingly, monthly values are computed by daily means, and months with less than `r Monthly_aggegation_N_lim` days are omitted. 
To create the daily and monthly seasonal values, we group the data based on the day of year number, and the corresponding calendar month respectively.
<!-- Seasonal -->
For the seasons of the year trends, we grouped the mean daily values by the corresponding season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, for each data set, we remove the natural occurring seasonal variation, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
<!-- SZA Deseasonalization -->
To estimate Sun's SZA contribution to the SDR brightening, the 1-minute data were aggregate in $1^\circ$ SZA bins, separated for instances before and after the local noon, and then was deseasonalized as above. 
Thus, for each dataset we have obtained the relative departure of the seasonal mean.

# Results

## Long-term SDR trends

We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
We present only the plot for the trend under all-sky conditions (Figure\ \ref{fig:trendALL}), as the plots for clear-sky condition and cloud-sky condition, are very similar.
In the studied time period, there is no significant break or change in the variability of the overall trends.
Other studies for the European region are reporting a change of the SDR slope, around the 1980, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].

<!-- \footnotesize -->
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
## no need all are 100%
LT$`Trend statistical signif. [%]` <- NULL

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". For all cases the statistical significance is above $99.9\\%$."))
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
## some vars to use later
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
```
<!-- \normalsize -->

```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily SDR relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line shows the long term trend for all-sky conditions.")}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
In our case, for all sky conditions we observe a period where the cumulative remain relative steady.
This indicates, that the anomaly was evenly distributed around the climatological mean.
Another distinct feature of the cumulative plots, is the difference, between the sky conditions, of the start of the sums sudden increase, for the latter part of the period.
This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.

```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
```

<!-- ```{r cumsumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics(c( -->
<!--     "./images/CumulativeDailyCumSumSTD-1.png",    # ALL sky -->
<!--     "./images/CumulativeDailyCumSumSTD-2.png",    # CLEAR sky -->
<!--     "./images/CumulativeDailyCumSumSTD-3.png"))   # Cloud sky -->
<!-- ``` -->

....

more about trend in relation to other trends
Aerosols and clouds

......

<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
<!-- knitr::include_graphics("./images/LongtermTrends-4.png") -->
<!-- ``` -->

<!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
<!-- knitr::include_graphics("./images/LongtermTrends-6.png") -->
<!-- ``` -->

## Effects of the solar zenith angle on SDR.

.... cite paper about low elevation SDR .....

The relative position of the Sun in the sky, seem to be a major drive of the attenuation of SDR, as it is affect

... Aerosols and clouds ...

We group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin

Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a seasonal distribution of SZA than in not presented here.

```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics("./images/SzaTrends-13.png")
knitr::include_graphics("./images/SzaTrends-17.png" )
knitr::include_graphics("./images/SzaTrends-21.png")
```

## Long term trends by season of year

Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
This can give us a better understanding of the annual variability of the trends.

```{r seasonalALL, echo=F, fig.cap="Trends by season for all condition. Displaying monthly means of daily means.", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether-1.png")
```

<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
<!-- ``` -->

<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
<!-- ``` -->

<!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-3.png") -->
<!-- ``` -->

<!-- \footnotesize -->
```{r trendseasontable, echo=F}
pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE)
pp <- pp[grep("SDR" ,pp$var), ]
pp$slope.sd           <- NULL
pp$slope.ConfInt_0.95 <- NULL
pp$slope.ConfInt_0.99 <- NULL
pp$var                <- NULL
rownames(pp) <- NULL
names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Trend statistical signif. [%]")

# panderOptions("table.alignment.default", "left")
panderOptions("table.alignment.default", "center")
pander(pp,
       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year.")
```
<!-- \normalsize -->

<!--
## Consistency of the trends

A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.

Figure\ \ref{fig:cumsum}

... different in aggregation ...
-->

<!--
```{r cumsumALL, echo=F, fig.cap="Running cumulative sum."}
knitr::include_graphics("./images/CumulativeDailyCumSum-1.png")
```
-->

<!--
Not showing this trend
```{r cumsumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
knitr::include_graphics("./images/CumulativeDailyCumSum-5.png")
```
-->

<!--
```{r cumsumCLOUD, echo=F, fig.cap="Donditions."}
knitr::include_graphics("./images/CumulativeDailyCumSum-9.png")
```
-->

# Conclusions

Our result for all-sky condition ($`r alltrend`\%/year$), reaffirm the previous results of @Bais2013 for the site.
The increase of this trend, shows that the phenomenon and probably the causes have been amplified.

.....
Also, the trend of $`r cloudtrend`\%/year$ for cloud-sky condition indicate the major part that clouds play on the
.....

.......

About:

- long terms trends
- seasonal trends
- effect of SZA
- effect of clouds

- Aerosols

similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
include in discussion

-------------------------------------------------
                   **END**
-------------------------------------------------
\newpage