### 2023-10-24 Article.Rmd
Commit 2023-10-24 13:09

---
title:         "Trends of SDR in Thessaloniki"
# header-includes:
author:
  - name:         Athanasios Natsis 
    email:        natsisphysicist@gmail.com
    affiliation:  Laboratory of Atmospheric Physics
#    footnote:     1
    correspondingauthor: true
  - name:         Alkiviadis Bais
    email:        abais@auth.gr 
    affiliation:  Laboratory of Atmospheric Physics
    correspondingauthor: false
address:
  - code:         Laboratory of Atmospheric Physics
    organization: Aristotle University of Thessaloniki
    addressline:  Campus Box 149
    city:         Thessaloniki
    # state:        State
    postcode:     54124
    country:      Greece
#  - code:         Another University
#    organization: Department
#    addressline:  A street 29
#    postcode:     2054 NX
#    city:         Manchester,
#    country:      The Netherlands
# footnote:
#   - code:         1
#     text:         "This is the first author footnote."
#   - code:         2
#     text:         "Another author footnote."
abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
  Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
  We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
  We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
keywords: 
  - GHI
  - SDR
  - Solar Brigthening/Dimming
journal:         "An awesome journal"
date:            "`r Sys.Date()`"
classoption:     preprint, 3p, authoryear
bibliography:    Mainreferences.bib
linenumbers:     false
numbersections:  true
header-includes:
   # - \usepackage{subfig}
   - \usepackage{subcaption}
output: 
  rticles::elsevier_article:
    keep_tex:         no
    citation_package: natbib
  bookdown::word_document2: default
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              no
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
# date:           "`r format(Sys.time(), '%F')`"
---

```{r echo=F, include=T, warning=FALSE, message=FALSE}
suppressMessages(
    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
)
## document configurations 
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "70%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
load("./data/common_data_14_2.Rda")
load("./data/Input_1_longterm_trends.Rda")
## number of measurements for each dataset
SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
## TSI data span info
TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
                    strip.white = TRUE,
                    sep = ";",
                    header = TRUE))
## variables we use in the analysis
source("./DHI_GHI_0_variables.R")

## create local citations file and links here
source("~/CODE/FUNCTIONS/R/gather_citations.R")
suppressMessages(
    gather_citations(
        sourcefile   = "./Article.Rmd",
        mainreposit  = "./Mainreferences.bib",
        localreposit = "./references.bib",
        folderlink   = "./References_links",
        workingdir   = "./"
    )
)
```

<!-- use SDR for GHI -->

# Introduction.

<!-- something about aerosols: bais mail -->
<!-- something about clouds: bais mail -->
<!--
Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
-->

The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Can also affect, solar and agricultural production, and all living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
<!-- Trends china -->
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
<!-- Trends Europe Brazil -->
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
<!-- Trends Global -->
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].

There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->

Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
<!-- goals -->
We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).

<!--
Aerosols interactions

#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China

Aerosols not only strengthen but also weaken the
growth of clouds with respect to their coverage and
optical thickness, depending on the levels of pollution
and the associated amounts of aerosols as postulated
in a conceptual framework (Wild 2009a, 2012, Yang
et al 2012).

...... cloud formation occurrences and constitution nuclei. cite .....
p>        "  Too hard to vim

#Wild2012 : simplify interaction nice graph

-->

# Observational data and methodology.

The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki 
($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.)
in the period from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS$Date)`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS$Date)`.
<!-- }. -->
The monitoring site is located near the city centre, and we expect to be affected by the urban environment. 
During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than $0.7\%$ relative to the initial calibration by the manufacturer. 
Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp & Zonen CHP-1 pyrheliometer, for the period
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
<!-- }. -->
Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets. 
There are four distinct steps in the creation of the dataset analysed here:
a)\ the acquisition of radiation measurements from the sensors,
b)\ the data quality check,
c)\ the identification of "clear sky" conditions from the radiometric data, and 
d)\ the aggregation of data and trend analysis.

<!--
The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.  These data points were excluded from the analysis, as
-->

<!-- Zerefos 
Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
loniki, the major city in northern Greece, lies along the exit of air
pollutants from Europe, in the path of the prevailing long-range
transport (LRT) trajectories, as we move from central and east-
ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
-->

<!-- a) acquisition -->
For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of $1\,\text{Hz}$.
The mean and the standard deviation of these samples are recorded every minute.
The measurements are corrected for the zero offset ("dark signal" in volts).
The "dark signal" is calculated by averaging all measurements recorded for a period of 
$3\,\text{h}$, 
before (morning) or after (evening) the Sun reaches an elevation angle of
$-10^\circ$.
The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.

<!-- b) Radiation data quality check -->
A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of 
Long and Shi\ [-@Long2008a; -@Long2006].
Thus, problematic recordings have been excluded from further processing. 
Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.

<!-- c) Clear sky identification -->
In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
To identify the clear-sky conditions we used a method proposed by
@Long2000
and by
@Reno2016,
which was adapted and configured for the site, as the authors suggest.

<!-- CS subjectivity -->
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.

```{r setup, echo=F, include=T, eval=TRUE}
## Load Clear Sky id v14.2 variables
MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
# export all Clear Sky variables
for (an in names(MS)) {
assign(an, MS[[an]])
}
## get TSI slope for reference
TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
           strip.white = TRUE,
           sep = ";",
           header = TRUE))
TSI <- TSI[grep("TSI", TSI$var), ]
TSI$slope.sd           <- NULL
TSI$slope.ConfInt_0.95 <- NULL
TSI$slope.ConfInt_0.99 <- NULL
TSI$var                <- NULL
TSI$N                  <- NULL
TSI$Obs                <- NULL
setorder(TSI, -slope.stat_sig)
rownames(TSI) <- NULL
names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
```

<!-- # Detection of clear periods in GHI measurements for SDR trends. -->

For completeness, we will provide below a brief overview of the 
<!--
[....Csid term is referenced above....]{.comment-start id="0" author="A N"
date="2023-07-15T20:00:52Z"} 
-->
clear sky identification algorithm (CSid),
<!--
[]{.comment-end id="0"} 
-->
along with the site specific thresholds.
To calculate the reference clear sky
$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
the radiation model of @Haurwitz1945, adjusted for our site with a
factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
by @Long2000 and @Reno2016.
The target of
the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by @Reno2012 and evaluated by @Reno2016.
We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
while the procedure, manages to characterize the majority of the data.
<!--[]{.comment-end id="3"}--> 
Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.

<!-- cost fuction -->
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation}

where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.

<!-- adjusted Haurwitz -->
\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = `r signif(alpha,3)` \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}

where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below.
A data point is flagged
as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
Each criterion was applied
for a running window of $`r nt`$ consecutive one-minute measurements, and
the characterization is assigned to the central value of the window.
Each parameter, was calculated both from the observations and the
reference clear sky model, for each comparison.
The allowable range of variation is defined by the
model-derived value of the parameter multiplied by a factor plus an
offset. 
<!--[.... .expanded the description ....]{.comment-start id="4"
author="A N" date="2023-07-15T22:48:13Z"}
-->
The factors and the offsets were
determined empirically, by manual inspecting each filters performance on
selected days, and adjusting them accordingly, during an iterative
process.
<!--[[]{.comment-end id="4"}]{.insertion author="A N"
date="2023-07-15T22:48:13Z"}
-->

<!-- 1. MeanVIP -->
<!-- ### Mean value of irradiance during the time period. -->

<!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq. \ref{eq:MeanVIP}).-->

a) Mean of the measured $\overline{\text{SDR}}_i$ (Eq. \ref{eq:MeanVIP}).
\begin{equation}
`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
< \overline{\text{SDR}}_i <
`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
\label{eq:MeanVIP}
\end{equation}

<!-- 2. MaxVIP -->
<!-- ### Max value of irradiance during the time period. -->

<!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}). -->

b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
\begin{equation}
`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`
< M_{\text{}i} <
`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`
\label{eq:MaxVIP}
\end{equation}

<!-- 3. VIL -->
<!-- ### Variability in irradiance by line length. -->
c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILeq}).
\begin{equation}
L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}
\begin{equation}
`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{\text{CSref},i} + `r offVIL_upl`
\label{eq:VILcr}
\end{equation}
where: $t_i$ is the time each SDR measurement has been measured

<!-- 4. VCT -->
<!-- ### Variance of Changes in the Time series. -->

<!-- Standard deviation of rate of change in irradiance. -->
d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Eq.\ \ref{eq:VCT1}).
<!-- \begin{gather}
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather} -->
\begin{gather}
  \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
  s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
\end{gather}
For this criterion, $\sigma_i$ should be below a certain threshold (Eq.\ \ref{eq:VCTcr}):
\begin{equation}
  \sigma_i < `r offVCT` \label{eq:VCTcr}
\end{equation}

<!-- 5. VSM -->
<!-- ### Variability in the Shape of the irradiance Measurements. -->

e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
\begin{gather}
  X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
  x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
  x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
\end{gather}
For this criterion, $X_i$ should be below a certain threshold (Eq.\ \ref{eq:VSMcr}):
\begin{equation}
  X_i < `r offVSM` \label{eq:VSMcr}
\end{equation}

<!-- ### Data selection -->
Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
On the latter instances, Sun is systematically, not visible by the instrument's location.
<!-- 1au -->
To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
<!-- TSI -->
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a <!--homogenized--> time series of satellite TSI observations.
The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
Where the initial daily values, were interpolated to match with the time step of our measurements. 
The final dataset contains
$`r SDR_N_All`$
one-minute measurements, of which, 
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
were identified as under clear-sky conditions and subsequently
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
as under cloud-sky conditions.
<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->

<!-- aggregation -->
In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
To preserve the representativeness of the data we used the following criteria: 
<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements, present and valid,
b) monthly values were computed from daily means
only when at least `r Monthly_aggegation_N_lim` days were available.
To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
<!-- Seasonal -->
For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.

# Results

## Long-term SDR trends

```{r echo=F, eval=TRUE}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
LT$N                  <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", 
                  "Trend p-value", 
                  "Sky conditions",
                  "Observations",
                  "Statistical signif. [%]")
## no need all are 100%
LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
# LT$`Trend statistical signif. [%]` <- NULL
## numeric precision
LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)

## some vars to use later
alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
## rename sky conditions
LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)

## order columns
LT1 <- LT[,c(
      "Sky conditions",
      "Trend [%/year]",
      "Trend p-value",
      "Observations"
      )]

## order columns
LT2 <- LT[,c(
      "Sky conditions",
      "Trend [%/year]",
      "Statistical signif. [%]",
      "Observations"
      )]

```

We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
We observe a positive trend for all-sky conditions 
($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009]. It is interesting to note, that for the observation period, the trend of the TSI is
$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.

<!-- \footnotesize -->
<!-- ```{r trendtable2, echo=F} -->
<!-- panderOptions("table.alignment.default", "center") -->
<!-- pander(LT1, -->
<!--        caption = paste0("\\label{tab:trendtable2}Trends in SDR daily mean for different sky conditions for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". For all cases the statistical significance is above $", LTminSSig, "\\,\\%$.")) -->
<!-- ``` -->
<!-- \normalsize -->

```{r trendtable, echo=F}
x_latex <- knitr::kable(LT2,
                        format   = "latex",
                        booktabs = T,
                        escape   = T,
                        format.args = list(digits = 4),
                        linesep  = "",
                        align    = "crr",
                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily mean for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
x_latex %>%
    kable_styling(
        latex_options = c("hold_position")
    )
        # position      =   "float_left"
```

```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line shows the long term linear trend.")}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019]. 
In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean. 
In our case, there is more complex evolution of the phenomenon.
Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).

In order to investigate more any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean. When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years. 
The pattern for the all sky and clear sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear- sky, have a high variability around zero. After 2016, the range of the variability is decreased at about one third of the prior period.
For cloud-sky conditions (Figure\ \ref{fig:cusumnotrend-3}) the period 1997-2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008. 
The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.

<!-- ```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics(c( -->
<!--     "./images/CumulativeDailyCuSum-1.png",    # ALL sky -->
<!--     "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky -->
<!--     "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky -->
<!-- ``` -->

```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
```

<!-- ```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics(c( -->
<!--     "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky -->
<!--     "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky -->
<!--     "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky -->
<!-- ``` -->

```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
```

<!-- ```{r cusumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics(c( -->
<!--     "./images/CumulativeDailyCuSumSTD-1.png",    # ALL sky -->
<!--     "./images/CumulativeDailyCuSumSTD-2.png",    # CLEAR sky -->
<!--     "./images/CumulativeDailyCuSumSTD-3.png"))   # Cloud sky -->
<!-- ``` -->

## Effects of the solar zenith angle on SDR.

The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
Although there are seasonal variation of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.

```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics("./images/SzaTrends-1.png")
knitr::include_graphics("./images/SzaTrends-4.png" )
knitr::include_graphics("./images/SzaTrends-7.png")
```

<!-- ```{r szatrendsmonthly, echo = FALSE, fig.cap='*** Montlhly *** Distribution of the SDR\'s long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
<!-- knitr::include_graphics("./images/SzaTrendsMonthly-1.png") -->
<!-- knitr::include_graphics("./images/SzaTrendsMonthly-4.png" ) -->
<!-- knitr::include_graphics("./images/SzaTrendsMonthly-7.png") -->
<!-- ``` -->

## Long term trends by season of year

```{r echo=F, eval=TRUE}
## Prepare table and data
spp <- read.csv("./figures/tbl_longterm_trends_season.dat",
               strip.white = TRUE,
               sep         = ";",
               header      = TRUE)
spp <- spp[grep("SDR" ,spp$var), ]
spp$slope.sd           <- NULL
spp$slope.ConfInt_0.95 <- NULL
spp$slope.ConfInt_0.99 <- NULL
spp$var                <- NULL
rownames(spp)          <- NULL
names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
## order by season
target    <- c("Winter", "Spring", "Summer", "Autumn")
new_order <- sapply(target, function(x,spp){which(spp$Season == x)}, spp = spp)
spp        <- spp[new_order,]
## order by sky condition
spp <- spp[order(spp$`Sky condition`),]

# ## put all sky first
# A  <- which(spp$`Sky condition` == "All sky cond.")
# B  <- which(spp$`Sky condition` != "All sky cond.")
# spp <- spp[c(A,B),]
## numeric precision
spp$`Trend [%/year]`          <- signif(spp$`Trend [%/year]`, 3)
spp$`Trend p-value`           <- signif(spp$`Trend p-value`,  3)
spp$`Statistical signif. [%]` <- signif(spp$`Statistical signif. [%]`,  3)

## rename sky conditions
spp$`Sky condition` <- sub(" cond.", "", spp$`Sky condition`)
```

Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
value similar to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
In spring and summer, the trends are much smaller and less statistical significant.
These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness. 
Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions. Each point is a monthly mean calculated from daily means. The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
```

<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for all-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
<!-- ``` -->

<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for clear-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
<!-- ``` -->

<!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
<!-- knitr::include_graphics("./images/SeasonalTrends-3.png") -->
<!-- ``` -->

<!-- \footnotesize -->

```{r trendseasontable, echo=F}
## arrange columns
spp <- spp[,c(
      "Sky condition",
      "Season",
      "Trend [%/year]", 
      # "Trend p-value", 
      "Statistical signif. [%]"
      )]

rownames(spp) <- NULL
# panderOptions("table.alignment.default", "left")
# panderOptions("table.alignment.default", "center")
# pander(spp,
#        justify = c('center', 'right', 'right', 'center'),
#        split.table = Inf,
#        caption = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")

## use kable

x_latex <- knitr::kable(spp,
                        format   = "latex",
                        booktabs = T,
                        escape   = T,
                        # format.args = list(digits = 3),
                        linesep  = "",
                        align    = "ccrr",
                        caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
x_latex %>%
    collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
    kable_styling(latex_options = c("striped",
                                    "hold_position")
                  ) 
# position = "float_left"

```

<!-- \normalsize -->

# Conclusions

We demonstrate that in the period 
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends). And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.

<!-- to conclutions? -->
In the CUSUMs for all-skies and clear-skies, we observed periods where the CUSUMs remain relative steady, 
<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after. This is an indication that the whole brightening phenomenon does not have a smooth development over time.

Our findings are in agreement with other researchers for the region.
Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki. 
Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.

<!--
similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
include in discussion
-->

# References {-}

<div id="refs"></div>

# Appendix {-}

```{r trendCLEAR, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily clear-sky SDR, relative to climatological values for", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line shows the long term linear trend for clear-sky conditions.")}
knitr::include_graphics("./images/LongtermTrends-5.png")
```

```{r trendCLOUD, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily cloud-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line shows the long term linear trend for cloud-sky conditions.")}
knitr::include_graphics("./images/LongtermTrends-8.png")
```

```{r SZAtrendSeason, echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")
```

<!-- ```{r echo = FALSE, fig.cap='*** Montly *** SDR trend ~ SZA', out.width='100%'} -->
<!-- knitr::include_graphics("./images/SzaTrendsSeasTogetherMonthly-2.png") -->
<!-- ``` -->