### 2023-11-01 Article.tex
Commit 2023-11-01 16:14

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles

\documentclass[An awesome
journal,,,moreauthors,pdftex]{Definitions/mdpi}

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Trends of SDR in Thessaloniki}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Trends of SDR in Thessaloniki}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Athanasios Natsis$^{}$, Alkiviadis Bais$^{}$}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais}

% MDPI internal command: Authors, for citation in the left column

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
}

% Contact information of the corresponding author
\corres{Correspondence: }

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. While monitoring the long term variability, there are
observations of upward and downward SDR trends on different locations
around the world for different time periods. Periods of positive tredns
are refered as brightening periods and with negative tredns as dimming
periods. We studied 29 years of CHP1 data from Thessaloniki, Greece,
under three sky conditions (clear sky, cloudy sky and all sky
conditions), applying a cloud sky identification algorithm. We found a
positive trend for all-sky and clear-sky conditions, and also,
investigated the consistency of those trends, the effect of the solar
zenith angle, and the variation of the trends for the seasons of the
year. We indentified that there are some anomalies in the long term SDR
trends, for all sky conditions.}

% Keywords
\keyword{GHISDRSolar Brigthening/Dimming}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction.}{%
\section{Introduction.}\label{introduction.}}

The shortwave downward solar irradiance (SDR) at Earth's surface play a
significant role, on its climate. Changes of the SDR can be related to
changes on Earth's energy budget, the mechanisms of climate change, and
water and carbon cycle \citep{Wild2009}. Can also affect, solar and
agricultural production, and all living organisms. Studies of SDR
variability, have identified some distinct SDR trends on different
regions of the world on different time periods. The term `brightening'
is generally used to describe periods of positive SDR trend, and
`dimming' for negative trend. There are many cases on the long term
records of irradiance, showing a systematic change of SDR's trend slope,
occurring roughly at the last decades of the 20th century. On multiple
station in China, a dimming period was reported until about 2000,
followed by a brightening period \citep{Yang2021}. A similar pattern was
identified, with the breaking point around 1980, for stations in Central
Europe \citep{Wild2021} and Brazil \citep{Yamasoe2021}. Also, on global
scale, an AI aided continental level spatial analysis, with data from
multiple station, reach similar conclusions for the above regions and
for the global trend \citep{Yuan2021}.

There is a consensus, among researchers, that the major factors of SDR
attenuation is the interaction of Sun radiation with atmospheric
aerosols and clouds. Those interactions, among other factors, have been
analysed with models \citep{Li2016, Samset2018}, showing the existence
of feedback mechanisms between the two. Similar finds, have been showed
in observational data \citep[ and references
therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.

Due to the variability of the phenomenon, and its contributing factors,
there is a constant need to investigate and monitor SDR, in different
sites, to estimate its magnitude, and its relation to the local
conditions. In this study, we examine the trends of SDR, with
ground-based measurements at Thessaloniki, Greece for the period 1993 to
2023, as derived from a CM-21 pyranometer. We reevaluated and extended
the dataset used by \citet{Bais2013}, applying a different algorithm for
the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
period 1993 to 2023, under different sky conditions (all-sky, clear-sky
and cloud-sky).

\hypertarget{observational-data-and-methodology.}{%
\section{Observational data and
methodology.}\label{observational-data-and-methodology.}}

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
2023-04-13. The monitoring site is located near the city centre, and we
expect to be affected by the urban environment. During the study period,
the pyranometer has been independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
stability of the instrument to within better than \(0.7\%\) relative to
the initial calibration by the manufacturer. Along with SDR, the direct
beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
we have performed a similar analysis to the DNI data the results are not
presented here, as they lack the appropriate statistical significance,
due to the sorter duration of the data. However, the DNI data were used
as auxiliary data, in the clear sky identification algorithm (CSid),
which is discussed later, for the selection of the appropriate
thresholds. It is noted that despite the capability of the CSid
algorithm to use the DNI as a characterization parameter, we haven't
utilized it here, to avoid any selection bias, due to unequal length of
the two datasets. There are four distinct steps in the creation of the
dataset analysed here: a)~the acquisition of radiation measurements from
the sensors, b)~the data quality check, c)~the identification of ``clear
sky'' conditions from the radiometric data, and d)~the aggregation of
data and trend analysis.

For the acquisition of radiometric data, the signal of the pyranometer
is sampled with a rate of \(1\,\text{Hz}\). The mean and the standard
deviation of these samples are recorded every minute. The measurements
are corrected for the zero offset (``dark signal'' in volts). The ``dark
signal'' is calculated by averaging all measurements recorded for a
period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
reaches an elevation angle of \(-10^\circ\). The signal is converted to
irradiance using a ramped value of the instrument's sensitivity between
calibrations.

A manual screening was performed, to remove inconsistent and erroneous
recordings that can occur stochastically or systematically, during the
continuous operation of the instruments. The manual screening is aided
by a radiation data quality assurance procedure, adjusted for the site,
which is based on the methods of Long and
Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
problematic recordings have been excluded from further processing.
Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the good
quality of the radiometric measurements used in this study.

In order to be able to estimate the effect of the sky condition on the
long term variability of SDR, we created three datasets, by
characterizing each one-minute measurement with a corresponding sky
condition (i.e., all-sky, clear-sky and cloudy-sky). To identify the
clear-sky conditions we used a method proposed by \citet{Long2000} and
by \citet{Reno2016}, which was adapted and configured for the site, as
the authors suggest.

We have to note, that the definition of what constitutes as clear or
cloudy sky, has some subjectivity, in any method of characterization. As
a result, the details of the definition are site specific, it relies on
a combination of thresholds and comparisons with ideal actinometric
models and statistical analysis on different signal metrics. The CSid
algorithm was calibrated with the main focus, to identify the presence
of clouds on the sky dome. Although the fine-tuning of the procedure,
few marginal cases exist, that have been identified manually as false
positive or false negative but cannot affect the final results of the
study.

For completeness, we will provide below a brief overview of the clear
sky identification algorithm (CSid), along with the site specific
thresholds. To calculate the reference clear sky
\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
optimization process, as described by \citet{Long2000} and
\citet{Reno2016}. The target of the optimization was the minimization of
a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
algorithmic function ``optimise'', which is an implementation based on
the work of \citet{Brent1973}, from the library ``stats'' of the R
programming language \citep{RCT2023}. The optimization and the selection
of the clear sky reference model, was performed on SDR observations for
the period 2016 - 2021. During the optimization, eight simple clear sky
radiation models were tested (Daneshyar-Paltridge-Proctor,
Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by \citet{Reno2012} and
evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
with the factor \(a = 0.965\) yields one of the lowest root mean squared
errors (RMSE), while the procedure, manages to characterize the majority
of the data. Thus, our clear sky reference is derived by the
Eq.~\ref{eq:ahau}.

\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation}

where: \(n\) is the total number of daylight records,
\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
by CSid, \(a\) is a hypothetical adjustment factor, and
\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
radiation models.

\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}

where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in
\(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle
(SZA).

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below. A data point is flagged
as ``clear-sky'' if all criteria are satisfied, otherwise it is
considered to be ``cloud-sky''. Each criterion was applied for a running
window of \(11\) consecutive one-minute measurements, and the
characterization is assigned to the central value of the window. Each
parameter, was calculated both from the observations and the reference
clear sky model, for each comparison. The allowable range of variation
is defined by the model-derived value of the parameter multiplied by a
factor plus an offset. The factors and the offsets were determined
empirically, by manual inspecting each filters performance on selected
days, and adjusting them accordingly, during an iterative process.

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
  \ref{eq:MeanVIP}). \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
  \label{eq:MeanVIP}
  \end{equation}
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75
  < M_{\text{}i} <
  1 \times M_{\text{CSref},i} + 75
  \label{eq:MaxVIP}
  \end{equation}
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  Length \(L_i\) of the sequential line segments, connecting the points
  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation} \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation} where: \(t_i\) is the time each SDR measurement has
  been measured
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\setcounter{enumi}{3}
\tightlist
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
  \(11\) sequential points, normalized by the mean
  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather} For this criterion, \(\sigma_i\) should be below a
  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\setcounter{enumi}{4}
\tightlist
\item
  Maximum difference \(X_i\) between the change in measured irradiance
  and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather} For this criterion, \(X_i\) should be below a certain
  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
    X_i < 7.5 \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

Due to a significant measurement uncertainty near the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all one-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
the Sun's intensity variability, using a time series of satellite TSI
observations. The TSI data we use are part of the ``NOAA Climate Data
Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. Where
the initial daily values, were interpolated to match with the time step
of our measurements. The final dataset contains \(6589967\) one-minute
measurements, of which, \(84.2\%\) were identified as under clear-sky
conditions and subsequently \(15.8\%\) as under cloud-sky conditions.

In order to investigate the SDR trends, we implemented an appropriate
aggregation scheme to the 1-minute data to derive a series in coarser
timescale. To preserve the representativeness of the data we used the
following criteria: a) for the daily mean values we accept days with
more than 50\% of the daytime measurements, present and valid, b)
monthly values were computed from daily means only when at least 20 days
were available. To create the daily and monthly climatological means, we
averaged the data based on the day of year and calendar month,
respectively. For the seasonal means we averaged the mean daily values
in each season (Winter: December - February, Spring: March - May, etc.).
Finally, each data set was deseasonalized by subtracting the
corresponding climatological annual cycle (daily or monthly) from the
actual data. To estimate SZA contribution to the SDR trends, the
one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
the morning and afternoon hours, and then were deseasonalized as
mentioned above.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{long-term-sdr-trends}{%
\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}

We calculated the linear SDR trends, from the departures of the mean
daily values from the seasonal values, and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend
for cloud-skies (\(-0.28\,\%/y\)). In the studied period, there is no
significant break or change in the variability of the time series. Other
studies for the European region report a change of the SDR slope, around
1980, a few years before the start of our records
\citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
for the observation period, the trend of the TSI is \(-2e-04\,\%/y\),
and thus we can not attribute any major effect on SDR trend to Solar
variability.

\begin{table}[!h]

\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\centering
\begin{tabular}[t]{crrc}
\toprule
Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
\midrule
All sky & 0.376 & 100.00 & 10256\\
Clear sky & 0.349 & 100.00 & 10256\\
Cloudy sky & -0.276 & 99.97 & 5067\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[h!]

{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}

}

\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
\end{figure}

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
\ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior,
plots of the cumulative sums (CUSUM) of the anomalies can reveal
different structures within all the three sky conditions
\citep{Regier2019}. In the cases of all-sky and clear-sky (Figures
\ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
macroscopic periods. A downward part from the start until about 2005, an
relative steady part until about 2016 and finally, a steep upward part
until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}), we
have a different sequence, it begins with a relative steady part until
1997, followed by a upward part until 2005, a long decline until 2020,
with a small positive slope until the present. For a uniform trend, we
would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
This would indicate that the anomalies are evenly distributed around the
climatological mean, and for a positive uniform trend, the first half to
be below and the other half above the climatological mean. In our case,
there is more complex evolution of the phenomenon. Another distinct
feature of the CUSUMs, is the different pattern of the cloudy-sky
dataset which peaks around the middle of the period. Although, it seems
to exist a complementary relation to the CUSUMs of the clear- and
all-sky cases, we can not assert that clouds are the main drive for this
relation, due to the great difference in the number of observational
data between the datasets (Table~\ref{tab:trendtable}).

\begin{figure}[h!]

{\centering \subfloat[All-skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }

}

\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
\end{figure}

In order to investigate more, the features within the trends, we created
another set of CUSUMs plots by subtracting the corresponding long term
trend from the SDR anomaly data, prior to the CUSUM calculation
(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
the CUSUMs diverge from zero can be interpreted as systematic variation
of SDR from the climatological mean. When the CUSUM is increasing, the
added values are above the climatological values of SDR trend and vice
versa. Overall, for all- and clear-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
anomalies diverge from the climatological value, each lasting for
several years. The pattern for the all-sky and clear-sky conditions is
very similar, suggesting prevalence in clear skies over Thessaloniki. It
is interesting that in the period 1993 - 2016 the anomalies of all-sky
and clear-sky, have a high variability around zero. After 2016, the
range of the variability is decreased at about one third of the prior
period. For cloud-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
dominated by positive CUSUMs suggesting a reduced effect of clouds on
SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
continuous decrease in the thickness of clouds, followed by a period of
rapid increase (within 3 years) in cloud optical thickness lasting up to
2008. The following stable period spans for about 15 years up to 2021
when CUSUMs start increasing again.

\begin{figure}[h!]

{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }

}

\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
\end{figure}

\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
\subsection{Effects of the solar zenith angle on
SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}

The solar zenith angle is a major driver of the attenuation of SDR
reaching the ground, as SDR is affected by the enhancement of the
radiation path in the atmosphere, especially in urban environment with
human activities emitting aerosols \citep{Wang2021}. In order to
estimate the effect of the SZA on the SDR trends, we grouped the anomaly
data in bins of \(1^\circ\) SZA, and calculated the overall trend for
each bin before noon and after noon (Figure~\ref{fig:szatrends}).
Although there are seasonal dependencies of the minimum SZA (see
Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not
be further examined here. For all-sky and clear-sky conditions the
brightening effect of SDR (positive trend) is stronger for large SZAs
(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
the morning and afternoon hours are more or less consistent with small
differences, which can be attributed to systematic diurnal variations of
aerosols, particularly during the warm period of the year
\citep{Wang2021}. For cloud-sky conditions
(Figure~\ref{fig:szatrends-3}), we can not discern any significant
dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
the trends for the period before noon are stronger and negative,
possibly associated with stronger attenuation by clouds under oblique
incidence angles.

\begin{figure}[h!]

{\centering \subfloat[All-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloud-sky conditions.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }

}

\caption{Distribution of the SDR's long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-trends-by-season-of-year}{%
\subsection{Long term trends by season of
year}\label{long-term-trends-by-season-of-year}}

Similar to the long term trends discussed above, we have calculated the
trend of the anomalies for the three different sky conditions, and for
each season of the year, using the corresponding mean monthly values
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
all-sky conditions the trend in SDR in winter is the largest
(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), value
similar to the long term trend) both statistically significant above the
\(99\,\%\) confidence level. In spring and summer, the trends are much
smaller and less statistical significant. These seasonal differences
indicate a possible relation of the trends to clouds during winter and
autumn. For clear-skies, the trend in winter is \(0.83\,\%/y\), larger
than for all-skies (\(0.69\,\%/y\)), which is another indication of
increasing cloud thickness. Moreover, the trends under clear- and
cloud-sky conditions are almost complementary to each other,
particularly for winter and autumn, where the signal is stronger. During
spring and summer the statistical significance is very low and the
actual trend too small for a meaningful comparison.

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}

}

\caption{SDR trends by season of the year for the three sky conditions. Each point is a monthly mean calculated from daily means. The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}\label{fig:seasonalALL}
\end{figure}

\begin{table}[!h]

\caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\centering
\begin{tabular}[t]{ccrr}
\toprule
Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\

 & Spring & 0.1450 & 81.8\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\

\multirow{-4}{*}{\centering\arraybackslash All sky} & Autumn & 0.4310 & 99.3\\
\cmidrule{1-4}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\

 & Spring & 0.0613 & 38.6\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\

\multirow{-4}{*}{\centering\arraybackslash Clear sky} & Autumn & 0.3670 & 97.2\\
\cmidrule{1-4}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\

 & Spring & -0.0991 & 37.2\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\

\multirow{-4}{*}{\centering\arraybackslash Cloudy sky} & Autumn & -0.4000 & 89.5\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

We demonstrate that in the period 1993 - 2023, there is a brightening
trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a previous
study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
to be \(0.33\,\%/y\). The increase of this trend, indicates that the
phenomenon and probably the causes, are still occurring. Accordingly, we
found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
further supports that the brightening can be independent of the cloud
occurrences, although, cloud formation can also attenuate SDR. We have
calculated the effect of TSI variability to be a negligible factor in
SDR variability. Unfortunately, there is no available data for this
period for the aerosols, in order to directly estimate their effect on
SDR, which is the most probable main cause. The attenuation of SDR by
aerosols over Europe have been proposed as major factor by
\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
SDR reaching the ground, and changes on their occurrences can effect the
net energy balance. Because we have no adequate data to investigate the
long term changes of cloud formation in the region, we can not infer if
the SDR trend we observe under cloud-sky is biased by changes in clouds
and local climate. The observed brightening of SDR over Thessaloniki; is
dependent on SZA (higher SZA, have stronger brightening trends), and
also, with the seasons of the year, where Winter and Autumn show a
significant strong trend, in contrast to Spring and Summer. Using CUSUMs
of the trends for all- and clear-skies, we observed periods where the
CUSUMs remain relative steady, with a steep decline previous to that,
and a steep incline after. This is an indication that the whole
brightening phenomenon does not follow a smooth development over time.

Our findings are in agreement with other researchers for the region.
Future observations from the collocated pyrheliometer, may provide us
with the means to further investigate the variability of Solar radiation
at ground level in Thessaloniki. Also, more data (cloud occurrence,
aerosols, atmospheric water vapour, etc.), will help us to determine the
extent of SDR attenuation by each factor.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\begin{figure}[h!]

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}

}

\caption{Anomaly (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

\begin{figure}[h!]

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}

}

\caption{Anomaly (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}

}

\caption{Seasonla distribution of the SDR's long term trends by SZA before and after noon. The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale. Bins with $p<0.005$ or with less than 85 data points may be missing from the view.}\label{fig:SZAtrendSeason}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}
\bibliography{Mainreferences.bib}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}