### 2023-12-19 MDPI_submission.tex

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles

\documentclass[applsci,article,submit,moreauthors,pdftex]{Definitions/mdpi}

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Trends from 30-year observations of downward solar irradiance in
Thessaloniki, Greece}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Trends from 30-year observations of downward solar
irradiance in Thessaloniki, Greece}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Athanasios
Natsis$^{1}$\href{https://orcid.org/0000-0002-5199-4119}
{\orcidicon}, Alkiviadis Bais$^{1,*}$, Charikleia Meleti$^{1}$}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais, Charikleia Meleti}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Natsis, A.; Bais, A.; Meleti, C.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.);
\href{mailto:meleti@auth.gr}{\nolinkurl{meleti@auth.gr}} (C.M.)\\
}

% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes, production and can affect all
living organisms. Observations of SDR at different locations around the
world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred as brightening periods
and of negative trends as dimming periods. Here we studied 30 years of
pyranometer data in Thessaloniki, Greece, under three types of sky
conditions (clear sky, cloudy sky and all sky). The clear-sky data were
identified by applying a cloud screening algorithm. We found a positive
trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
conditions. We have also investigated the consistency of these trends,
their seasonal variability, and the effect of the solar zenith angle. We
have found that for all-sky and clear-sky conditions the SDR trend is
positive in winter (\(0.7\) and \(0.8\,\%/\text{year}\), respectively)
and autumn (\(~0.4\,\%/\text{year}\)), while under cloudy skies the
trend is negative (\(-0.9\,\%/\text{year}\) in winter and
\(-0.4\,\%/\text{year}\) in autumn). In spring and summer the trend is
very close to zero, irrespective of sky conditions. The SDR trend is
increasing with increasing solar zenith angle, except under cloudy skies
where the trend is highly variable and close to zero. Finally, we
identified some anomalies in the long term SDR trends for all sky
conditions by examining the cumulative sums of monthly anomalies from
the climatological mean.}

% Keywords
\keyword{GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols;
clouds.}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{subcaption}
\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The shortwave downward solar irradiance (SDR) at Earth's surface plays a
significant role, on its climate. Changes of the SDR can be related to
changes on Earth's energy budget, the mechanisms of climate change, and
water and carbon cycles \citep{Wild2009}. It can also affect solar and
agricultural production, and all living organisms. Studies of SDR
variability, have identified some distinct SDR trends on different
regions of the world on different time periods. The term `brightening'
is generally used to describe periods of positive SDR trend, and
`dimming' for negative trend \citep{Wild2009}. There are many cases in
the long term records of irradiance, showing a systematic change in the
magnitude of the trend, occurring roughly in the last decades of the
20th century. On multiple stations in China, a dimming period was
reported until about 2000, followed by a brightening period
\citep{Yang2021}. A similar pattern was identified, with the breaking
point around 1980, for stations in Central Europe \citep{Wild2021} and
Brazil \citep{Yamasoe2021}. On global scale, an artificial Intelligence
aided spatial analysis on continental level with data from multiple
stations reach similar conclusions for these regions and for the global
trend \citep{Yuan2021}.

There is a consensus among researchers that the major factor affecting
the variability of SDR attenuation is the interactions of solar
radiation with atmospheric aerosols and clouds. Those interactions,
among other factors, have been analysed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar findings have been shown in observational data
\citep[ and references
therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.

Due to the significant spatial and temporal variability of the trends,
and the contributing factors, there is a constant need to monitor and
investigate SDR in different sites in order to estimate the degree of
variability, and its relation to the local conditions. In this study, we
examine the trends of SDR, using ground-based measurements at
Thessaloniki, Greece, for the period 1993 to 2023, as derived from a
CM-21 pyranometer. We reevaluated and extended the dataset used by
\citet{Bais2013}, we applied a different algorithm for the
identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
period of study, under different sky conditions (all-sky, clear-sky and
cloud-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

\hypertarget{data-and-methodology}{%
\section{Data and methodology}\label{data-and-methodology}}

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) since 1993. Here we use data for
the period from 1993-04-13 to 2023-04-13. The monitoring site is located
near the city centre thus we expect that measurements are affected by
the urban environment. During the study period, the pyranometer has been
independently calibrated three times at the Meteorologisches
Observatorium Lindenberg, DWD, verifying that stability of the
instrument's sensitivity is better than \(0.7\%\) relative to the
initial calibration by the manufacturer. Along with SDR, the direct beam
radiation (DNI) is also measured with a collocated Kipp \& Zonen CHP-1
pyrheliometer since 2016-04-01. The DNI data were used as auxiliary data
to support the selection of appropriate thresholds in the clear sky
identification algorithm (CSid), which is discussed later. It is noted
that the limited dataset of DNI was not used for the identification of
clear sky cases in the entire SDR series to avoid any selection bias due
to unequal length of the two datasets. There are four distinct steps in
the creation of the dataset analyzed here: a)~the acquisition of
radiation measurements from the sensors, b)~the data quality check,
c)~the identification of ``clear sky'' conditions from the radiometric
data, and d)~the aggregation of data and trend analysis.

For the acquisition of radiometric data, the signal of the pyranometer
is sampled at a rate of \(1\,\text{Hz}\). The mean and the standard
deviation of these samples are calculated and recorded every minute. The
measurements are corrected for the zero offset (``dark signal'' in
volts), which is calculated by averaging all measurements recorded for a
period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
reaches an elevation angle of \(-10^\circ\). The signal is converted to
irradiance using a ramped value of the instrument's sensitivity between
subsequent calibrations.

A manual screening was performed, to remove inconsistent and erroneous
recordings that can occur stochastically or systematically, during the
continuous operation of the instruments. The manual screening is aided
by a radiation data quality assurance procedure, adjusted for the site,
which is based on the methods of Long and
Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
problematic recordings have been excluded from further processing.
Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality
of the radiation measurements used in this study.

Due to the significant measurement uncertainty when the Sun is near the
horizon, we have excluded all measurements with SZA greater than
\(85^\circ\). Moreover, due to obstructions around the site (hills and
buildings) which block the direct irradiance, we excluded data with
azimuth angle in the range \(58^{\circ}\) - \(120^{\circ}\) and with SZA
greater than \(78^{\circ}\). To make the measurements comparable
throughout the dataset, we adjusted all one-minute data to the mean Sun
- Earth distance. Subsequently, we adjusted all measurements to the
Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate
for the Sun's intensity variability, using a time series of satellite
TSI observations. The TSI data we used are part of the ``NOAA Climate
Data Record of Total Solar Irradiance'' dataset \citep{Coddington2005}.
The initial daily values of this dataset were interpolated to match the
time step of our measurements.

In order to estimate the effect of the sky conditions on the long term
variability of SDR, we created three datasets by characterizing each
one-minute measurement with a corresponding sky-condition flag (i.e.,
all-sky, clear-sky and cloudy-sky). To identify the clear-cases we used
a method proposed by \citet{Long2000} and by \citet{Reno2016}, which
were adapted and configured for the site of Thessaloniki. We note that
all methods have some subjectivity in the definition of clear or cloudy
sky cases. As a result, the details of the definition are site specific
and they rely on a combination of thresholds and comparisons with ideal
radiation models and statistical analysis of different signal metrics.
The CSid algorithm was calibrated with the main focus, to identify the
presence of clouds. Despite the fine-tuning of the procedure, in a few
marginal cases false positive or false negative results were identified
by manual inspection. However, due to their small number, they cannot
affect the final results of the study. For completeness, we provide
below a brief overview of the clear-sky identification algorithm (CSid),
along with the site specific thresholds.

\hypertarget{the-clear-sky-identification-algorithm}{%
\subsection{The clear sky identification
algorithm}\label{the-clear-sky-identification-algorithm}}

To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\) we used
the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
\citet{Haurwitz1945} (Eq.~\ref{eq:hau}), adjusted for our site:
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
\end{equation} where \(\theta\) is the solar zenith angle (SZA).

The adjustment was made with a factor \(a\) (Eq.~\ref{eq:ahau}), which
was estimated through an iterative optimization process, as described by
\citet{Long2000} and \citet{Reno2016}. The target of the optimization
was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and was
accomplished with the algorithmic function ``optimise'', which is an
implementation based on the work of \citet{Brent1973}, from the library
``stats'' of the R programming language \citep{RCT2023}.
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation} where: \(n\) is the total number of daylight data,
\(\text{SDR}_{\text{CSid},i}\) are the data identified as clear sky by
CSid, \(a\) is a site-specific adjustment factor, and
\(\text{SDR}_{\text{testCSref},i}\) is the SDR derived by any of the
tested clear-sky radiation models.

The optimization and the selection of the clear sky reference model, was
performed on SDR observations for the period 2016 - 2021. During the
optimization, eight simple clear sky radiation models were tested
(namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz,
Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and
Ineichen-Perez), with a wide range of factors. These models are
described in more details by \citet{Reno2012} and evaluated by
\citet{Reno2016}. We found, that Haurwitz's model, adjusted with the
factor \(a = 0.965\) yields one of the lowest root mean squared errors
(RMSE), while the procedure manages to characterize successfully the
majority of the data. Thus, our clear sky reference is derived by the
Eq.~\ref{eq:ahau}: \begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below. A data point is flagged
as ``clear-sky'' if all criteria are satisfied; otherwise it is
considered as ``cloud-sky''. Each criterion was applied for a running
window of \(11\) consecutive one-minute measurements, and the
characterization was assigned to the central datum of the window. Each
parameter, was calculated both from the observations and the reference
clear sky model, for each comparison. The allowable range of variation
is defined by the model-derived value of the parameter multiplied by a
factor plus an offset. The factors and the offsets were determined
empirically, by manually inspecting each filter's performance on
selected days and adjusting them accordingly during an iterative
process. The criteria are listed below, together with the range of
values within which the respective parameter should fall in order to
raise the clear-sky flag:

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
  \ref{eq:MeanVIP}). \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75\,Wm^{-2}
  < M_{i} <
  1 \times M_{\text{CSref},i} + 75\,Wm^{-2}
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points
  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation} \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation} where: \(t_i\) is the time stamp of each SDR
  measurement.
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
  \(11\) sequential points, normalized by the mean
  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather} For this criterion, \(\sigma_i\) should be below a
  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance
  and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather} For this criterion, \(X_i\) should be below a certain
  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
    X_i < 7.5\,Wm^{-2} \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

In the final dataset \(23.3\%\) of the 1-minute data were identified as
under clear-sky conditions and \(43\%\) as under cloud-sky conditions.

\hypertarget{aggregationstatistical}{%
\subsection{Aggregation of data and statistical
approach}\label{aggregationstatistical}}

In order to investigate the SDR trends which are the main focus of the
study, we implemented an appropriate aggregation scheme to the
one-minute data to derive a series in coarser time-scales. To preserve
the representativeness of the data we used the following criteria: a) we
excluded all days with less than 50\% of the expected daytime
measurements, b) daily means for the clear-sky and cloudy-sky datasets
were calculated only for days with more than 60\% of the expected
daytime measurements identified as clear or cloudy respectively, c)
monthly means were computed from daily means, for all-skies dataset
means were computed only when at least 20 days of the month were
available. Seasonal means were derived by averaging the monthly mean
values in each season (winter: December - February, spring: March - May,
etc.). The daily and monthly climatological means were derived by
averaging the data for each day of year and calendar month,
respectively. The daily and monthly datasets were deseasonalized by
subtracting the corresponding climatological annual cycle (daily or
monthly) from the actual data. Finally, to estimate the SZA effect on
the SDR trends, the one-minute data were aggregated in \(1^{\circ}\) SZA
bins, separately for the morning and afternoon hours.

The linear trends were calculated using an first order autoregressive
model with lag of 1 day with fitting method of the ``maximum
likelihood'' \citep{Gardner1980, Jones1980}, by implementing the
function ``arima'' from the library ``stats'' of the R programming
language \citep{RCT2023}. The trends were reported together with the
\(2\sigma\) errors (Table~\ref{tab:trendtable}).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{long-term-sdr-trends}{%
\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}

We calculated the linear trends of SDR, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In the studied
period, there is no significant break or change in the variability
pattern of the time series. The linear trends in all three datasets are
positive and around \(0.4\,\%/y\) for all-sky and cloudy-sky conditions,
while for clear-skies the trend is much smaller
(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
taking into account the autocorrelation of the time series and all three
are statistically significant at least at the \(95\,\%\) confidence
level, as they are larger than the corresponding \(2\sigma\) errors. The
clear-sky trend is very small suggesting a small effect from aerosols
and water vapor which are the dominant factors of the SDR variability.
In contrast the large positive trend of SDR under cloudy skies can be
attributed to reduction in cloud cover and/or cloud optical depth. Lack
of continuous cloud observations that could support these findings does
not allow drawing firm conclusions. However, there are indications that
the total cloud-cover as inferred from the ERA5 analysis for the grid
point of Thessaloniki is decreasing over the period of study.

The all-sky trend is similar to the one reported in \citet{Bais2013}
from a ten-year shorter dataset suggesting that the tendency of SDR in
Thessaloniki is systematic. Other studies for the European region
reported a change in the SDR trend around 1980 from negative to positive
with comparable magnitude \citep{Wild2021, Yuan2021, Ohmura2009}, well
before the start of our records. However, the trends reported here for
the three datasets are in accordance with the widely accepted solar
radiation brightening over Europe. For the period of our observations
the trend in the TSI is negligible (\(-0.00022\,\%/y\)), and thus we
cannot attribute any significant effect on the SDR trend to solar
variability.

\begin{table}[H]

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
\midrule
All skies & 0.380 & 0.120 & 0.091 & 1.5 & 10251\\
Clear skies & 0.097 & 0.033 & 0.140 & 0.5 & 2684\\
Cloudy skies & 0.410 & 0.180 & 0.081 & 1.2 & 4937\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}[h!]

{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}

}

\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period 1993 -- 2023. The black line is the long term linear trend.}\label{fig:trendALL}
\end{figure}

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR},
\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour,
plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
can reveal different structures in the records of all three sky
conditions. For time series with a uniform trend, we would expect the
CUSUMs of the anomalies to have a symmetric `V' shape centered around
the middle of the series. This would indicate that the anomalies are
evenly distributed around the climatological mean, and for a positive
uniform trend, the first half is below and the second half above the
climatological mean. In our case, there is a more complex evolution of
the anomalies. For all-skies (Figure~\ref{fig:cusummonth-1}), we observe
three rather distinct periods: a) a downward part between the start of
the datasets and about 2000, denoting that all anomalies are negative
thus below the climatology; b) a relatively steady part lasting for
almost 20 years suggesting little variability in SDR anomalies; and c) a
steep upward part to present indicating anomalies above the climatology.
The CUSUMs for cloudy-skies (Figure~\ref{fig:cusummonth-3}), show a
similar behavior with some short-term differences that do not change the
overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
monotonic downward tendency is evident until 2004, suggesting that the
anomalies are all negative. After 2004 the anomalies turn to positive at
a fast rate for about five years and at a slower rate thereafter.

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
        \subfloat[All skies.\label{fig:cusummonth-1}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
        \subfloat[Clear skies.\label{fig:cusummonth-2}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
        \subfloat[Cloudy skies.\label{fig:cusummonth-3}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
        }
\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
\end{adjustwidth}
\end{figure}

In order to unveil further the features of the variability of the three
datasets, Figure~\ref{fig:cusumnotrendmonthly} presents another set of
CUSUM plots using anomalies after the long-term linear trend is removed.
With this approach, periods when the CUSUMs diverge from zero can be
interpreted as a systematic variation of SDR from the climatological
mean. When the CUSUM is increasing, the anomalies values are above the
climatology and vice versa. Overall, for all- and cloudy-sky conditions
(Figures~\ref{fig:cusumnotrendmonthly-1}
and~\ref{fig:cusumnotrendmonthly-3}) we observe periods with anomalies
diverging from the climatological values, each lasting for several
years. These fluctuations are probably within the natural variability
and no distinct changes are identified. The pattern in both datasets is
similar, suggesting prevalence in cloudy skies over Thessaloniki. For
clear skies (Figure~\ref{fig:cusumnotrendmonthly-2}) the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. At that year there a
change in the rate of decrease in aerosol optical depth as illustrated
in Figure 7 of \citet{Siomos2020}. This abrupt change in CUSUMs lasts
until about 2010 when the anomalies become again variable.

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
            \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
            \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
        }
        \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
\end{adjustwidth}
\end{figure}

\hypertarget{effects-of-the-solar-zenith-angle-on-sdr}{%
\subsection{Effects of the solar zenith angle on
SDR}\label{effects-of-the-solar-zenith-angle-on-sdr}}

The solar zenith angle is a major factor affecting the SDR, since
increases in SZA leads to enhancement of the radiation path in the
atmosphere, especially in urban environments with human activities
emitting aerosols \citep{Wang2021}. In order to estimate the effect of
SZA on the SDR trends, we grouped the data in bins of \(1^\circ\) SZA,
and calculated the overall trend for each bin, separately for the daily
periods before noon and after noon (Figure~\ref{fig:szatrends}).
Although there are seasonal dependencies of the minimum SZA (see
Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies are not
discussed further.

For all-sky conditions the brightening effect of SDR (positive trend)
increases with SZAs (Figures~\ref{fig:szatrends-1}) ranging from about
\(0.1\,\%/y\) to about \(0.7\,\%/y\) for the statistically significant
trends. The trends in the morning and afternoon hours are more or less
consistent with small differences at small SZAs which can be attributed
to effects on clear sky SDR from systematic diurnal patterns of aerosols
during the warm period of the year, consistently with the results
reported for China by \citet{Wang2021}. Note that SZAs less than
\(25^\circ\) can only occur during the warm period of the year around
noon when clear skies are more frequent. The increasing of the trend
with SZA is likely caused by the increased attenuation of SDR with SZA.
The effect is larger when aerosol and/or cloud layers are optically
thicker, therefore, decreases in aerosol and clouds through the study
period will result in larger positive trends of SDR at larger SZAs.

Under clear skies (Figures~\ref{fig:szatrends-2}), the trends are
smaller and less variable, ranging between \(0.1\) and \(0.15\,\%/y\) up
to \(77^\circ\) SZA. At higher SZAs and in the afternoon hours there a
sharp increase in the trend up to \(0.3\,\%/y\), which may have been
caused by the long path length of radiation through the atmosphere as
discussed above for the all-sky conditions. The small

For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
discern any significant dependence of the SDR trend with SZA as the
variability of irradiance is dominated by the cloud effects. For SZAs
\(25^\circ\) -- \(75^\circ\), the trends are variable and for most SZAs
insignificant ranging between \(0\) and \(0.2\,\%/y\). The smooth
increase of the trend at SZAs larger than \(\sim 75^\circ\) is probably
associated with stronger attenuation by clouds under oblique incidence
angles which results in smaller variability.

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
            \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }
        \caption{Long term trends of SDR as a function of SZA separately from morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
    \end{adjustwidth}
\end{figure}

\hypertarget{long-term-trends-by-season}{%
\subsection{Long term trends by
season}\label{long-term-trends-by-season}}

Similarly to the long term trends from daily means of SDR discussed
above, we have calculated the trend for the three sky conditions and for
each season of the year, using the corresponding mean monthly anomalies
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
all-sky conditions the trend in SDR in winter is the largest
(\(0.7\,\%/y\)), followed by the trend in autumn (\(0.42\,\%/y\), a
value close to the long term trend) both statistically significant at
the \(95\,\%\) confidence level. In spring and summer, the trends are
much smaller and of lesser statistical significance. These seasonal
differences indicate a possible relation of the trends in SDR to trends
in clouds during winter and autumn. For clear-skies, the trend in winter
is \(0.4\,\%/y\), and is associated with the decreasing trend in aerosol
optical depth \citep{Siomos2020}, and almost half of that for all-skies,
which is another indication of a decreasing trend in cloud optical
thickness. In other seasons the clear-sky trend is still very small
(below \(0.1\,\%/y\)). Finally, for cloud-skies the winter trend is the
largest (\(0.76\,\%/y\)) and greater than for all-skies, followed by a
much smaller trend in autumn (\(0.19\,\%/y\)).

The trends under clear- and cloudy-sky conditions are of the same
direction and it would be expected that their sum is similar to the
all-sky trend. This does not happen, especially for winter, likely due
to the way the monthly means for clear and cloudy skies were calculated.
Daily means were calculated only when at least \(60\,\%\) of the clear-
or cloudy-sky data were available (see sect.
\ref{aggregationstatistical}).

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
        }
        \caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
    \end{adjustwidth}
\end{figure}

\begin{table}[!h]

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky condition & Season & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend p-value\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.70} & \cellcolor{gray!6}{0.43} & \cellcolor{gray!6}{0.54} & \cellcolor{gray!6}{0.003}\\

 & Spring & 0.11 & 0.24 & 0.17 & 0.371\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.11} & \cellcolor{gray!6}{0.15} & \cellcolor{gray!6}{0.25} & \cellcolor{gray!6}{0.175}\\

\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.42 & 0.30 & 0.47 & 0.009\\
\cmidrule{1-6}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.20} & \cellcolor{gray!6}{0.60} & \cellcolor{gray!6}{0.001}\\

 & Spring & 0.06 & 0.17 & 0.13 & 0.497\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.05} & \cellcolor{gray!6}{0.06} & \cellcolor{gray!6}{-0.30} & \cellcolor{gray!6}{0.106}\\

\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.05 & 0.12 & 0.17 & 0.366\\
\cmidrule{1-6}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.76} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.59} & \cellcolor{gray!6}{0.001}\\

 & Spring & 0.06 & 0.23 & 0.10 & 0.593\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.08} & \cellcolor{gray!6}{0.27} & \cellcolor{gray!6}{-0.11} & \cellcolor{gray!6}{0.560}\\

\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.19 & 0.43 & 0.16 & 0.384\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
Greece (1993 -- 2023) aiming to identify the long term variability of
solar irradiance under different sky conditions. Under all-sky
conditions there is a positive trend in SDR of \(0.38\,\%/y\)
(brightening). A previous study \citep{Bais2013} for the period 1993 --
2011 reported also a positive trend, of \(0.33\,\%/y\). The slight
increase of this trend indicates that the brightening of SDR continues
and is probably caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. The attenuation
of SDR by aerosols over Europe has been proposed as major factor by
\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
also decreasing during this period. As long term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative conclusions.

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend of \(0.7\) and \(0.76\,\%/y\) for all- and cloudy-skies,
respectively, in contrast to spring and summer. The trends for autumn
are also significant but smaller ( \(0.42\) and \(0.19\,\%/y\) for all-
and cloudy-skies, respectively). Our findings are in agreement with
other studies for the region. The trend for clear skies is largest in
winter (\(0.4\)) and negligible in spring, summer and autumn.

Using the CUSUMs of the monthly departures for all- and cloudy-skies, we
observed a 20 year period starting around 2000 where the CUSUMs remain
relatively stable, with a steep decline before and a steep increase
after. The rather smooth course of the CUSUMs suggests that no important
change in the SDR pattern has occurred in the entire record.

Continued observations with a collocated pyrheliometer, which started in
2016, will allow us to further investigate the variability of solar
radiation at ground level in Thessaloniki. Also, additional data of
cloudiness, aerosols, atmospheric water vapour, etc., will allow better
attribution and quantification of the effects of these factors on SRD.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\funding{This research received no external funding.}

\dataavailability{Data as daily sums are available through the WRDC
database \url{http://wrdc.mgo.rssi.ru}. One minute data are available on
request from the corresponding author. The data are not publicly
available for protection against unmonitored commercial use.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
DNI & Direct beam/normal irradiance \\
CSid & Clear sky identification algorithm \\
CUSUM & Cumulative sum \\
SDR & Solar downward radiation \\
SZA & Solar zenith angle \\
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\input{"appendix.tex"}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}
\bibliography{manualreferences.bib}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}